---
title: "TRM_University_PrepareData"
output: html_notebook
---

```{r setup, include=FALSE}
packages_vector <- c("tidyverse",
                     "readxl",
                     "sf",
                     "foreign")
need_to_install <- packages_vector[!(packages_vector %in% installed.packages()[,"Package"])]
if (length(need_to_install)) install.packages(need_to_install)
for (package in packages_vector){
  library(package, character.only = TRUE)
}
knitr::opts_chunk$set(echo = TRUE)

```


##FILE LOCATIONS
```{r remote-io}

private_dir <- "../data/input/_PRIVATE/"
data_dir<-"../data/input/"
prepared_data_dir<-"../data/input/university/"

#surveys
NCSUsurvey_name<-paste0(private_dir,"2013_NCSU_Data_Raw_All_2020-05-14_ITRE.xls")
ODU_name<-paste0(private_dir,"ODU1_adjustedNtrips.sav") 
UVA_name<-paste0(private_dir,"UVA1_adjustedNtrips.sav")
VCU_name<-paste0(private_dir,"VCU1_adjustedNtrips.sav")
VT_name<-paste0(private_dir,"VT1_adjustedNtrips.sav")
ODUw_name<-paste0(private_dir,"snVDOTCORRADINO_ODU_2010_with weight.sav") 
VTw_name<-paste0(private_dir,"snVDOTCORRADINO_VT_2010_with weight.sav")
dictUVA_name<-paste0(private_dir,"datadictionary.xlsx")

#zonal
taz_shape_file_name<-paste0(data_dir,"tazs/master_tazs.shp")
#adjacency_filename <-"Adjacent.csv"

#socioeconomics
se_filename<-paste0(data_dir,"se_data/se_2016.csv") # replace with master_tazs.bin converted to csv

```


##READ FILES
A survey of NCSU students conducted by the Institute for Transportation Research and Education (ITRE) for the North Carolina Department of Transportation (NCDOT) in 2014 serves as the basis for the estimation of trip generation rates for students from the Triangle Region's 4 main universities: NCSU,UNC-CH,Duke and NCCU. Findings from surveys of students from 4 universities in Virginia, as summarized in Khattak et al (2012), were used to verify the rates.
```{r}
#NCSU surveys
NCSU_Person_df<-read_excel(NCSUsurvey_name,"Person")
NCSU_Trip_df<-read_excel(NCSUsurvey_name,"Trip")
NCSU_Place_df<-read_excel(NCSUsurvey_name,"Place")
NCSU_DDtemp_df<-read_excel(NCSUsurvey_name,"Dictionary")
NCSU_DataDictionary_df<-NCSU_DDtemp_df[-c(2:3),]


# join person and trip data (place data already included in trip data)
NCSU_All_df<- NCSU_Person_df %>% 
full_join(NCSU_Trip_df, by="Person_ID")


#Virginia surveys
ODU_df<-ODUw_name %>% 
read.spss(to.data.frame = TRUE)
VT_df<-VTw_name %>% 
read.spss(to.data.frame = TRUE)
ODU_0_df<-ODU_name %>%  
read.spss(to.data.frame = TRUE)
VCU_0_df<-VCU_name %>%  
read.spss(to.data.frame = TRUE)
UVA_0_df<-UVA_name %>%  
read.spss(to.data.frame = TRUE)
VT_0_df<-VT_name %>%  
read.spss(to.data.frame = TRUE)
UVA_dd_df<-read_excel(dictUVA_name,"UVA1")


#Zonal
taz_sf <- st_read(taz_shape_file_name, stringsAsFactors = FALSE)
socioecon <-read_csv(se_filename)   
#adjacency<-read_csv(adjacency_filename)
#adjacency$TAZ<-as.character(adjacency$TAZ)

```
## PREPARE DATA
### TRIP-LEVEL DATA
#### step 1 - assign trip locations to TAZs
Assign start and end locations from NCSU survey to TAZs
```{r}
# join trip locations to tazs
PLANAR_EPSG <- 3857
LATLON_EPSG <- 4326
taz_join_sf <- st_transform(taz_sf, PLANAR_EPSG)

orig_tazs_df <- NCSU_All_df %>% 
  filter(Finished==1) %>% 
  select(Start_PlaceID,Start_Lat,Start_Long) %>%
  filter(!is.na(Start_Lat)) %>% 
  st_as_sf(., coords = c("Start_Long", "Start_Lat"),crs=LATLON_EPSG) %>%
  st_transform(., PLANAR_EPSG) %>%
  st_join(., taz_join_sf, join = st_intersects)

dest_tazs_df <- NCSU_All_df %>% 
  filter(Finished==1) %>% 
  select(End_PlaceID,End_Lat,End_Long)%>% 
  filter(!is.na(End_Lat))%>% 
  st_as_sf(., coords = c("End_Long", "End_Lat"),crs=LATLON_EPSG) %>%
  st_transform(., PLANAR_EPSG) %>%
  st_join(., taz_join_sf, join = st_intersects)

```

#### step 2 - recode/create new variables 
Recode and add new variables as follows:

##### Origin TAZ and Destination TAZ 
Add ID_o and ID_d based on Step 1
##### Trip Purpose
The model will include the following trip purposes segmented by on-campus and off-campus students:
•	Home-Based-Campus (UHC)
•	Home-Based-Other (UHO)
•	Campus-Based-Other (UCO)
•	On-Campus (UC1)
•	Inter-Campus (UCC)
•	University student Other-Other (UOO)
##### Production - Attraction
The Production trip end is home, campus (if home is not a trip end), or origin (if neither home or campus are a trip end).
##### Distance
A variable named "Distance" is used to indicate respondents who live in TAZs that are adjacent to the NCSU campus TAZs or TAZs that are adjacent to the adjacent TAZs (Distance==1)
##### Mode
For respondents who reported multiple modes, Mode_1 was assumed to be the primary mode.  Drive and car passenger are combined into a single mode named car. 

```{r}

NCSUtemp_df<- NCSU_All_df %>% 
  left_join(orig_tazs_df, by="Start_PlaceID")%>% 
  left_join(dest_tazs_df,by="End_PlaceID")%>%
  mutate(
    #ID_o=as.character(ID.x),
    #ID_d=as.character(ID.y),
    Trip_Purpose=case_when(
    Start_PlaceType=="Home" & End_PlaceType=="Home" ~ "99",
    Start_PlaceType=="Home" & End_PlaceType=="North Carolina State University" ~ "UHC", 
    End_PlaceType=="Home" & Start_PlaceType=="North Carolina State University" ~ "UHC",
    Start_PlaceType=="Home" & End_PlaceType!="North Carolina State University" ~ "UHO",
    End_PlaceType=="Home" & Start_PlaceType!="North Carolina State University" ~ "UHO",
    Start_PlaceType!="Home" & Start_PlaceType != "North Carolina State University"  & End_PlaceType=="North Carolina State University" ~ "UCO",
    End_PlaceType!="Home" & End_PlaceType != "North Carolina State University" & Start_PlaceType=="North Carolina State University" ~ "UCO",
    Start_PlaceType=="North Carolina State University" & End_PlaceType=="North Carolina State University" ~ "UC1",
    Start_PlaceType=="Other" & End_PlaceType=="Other" ~ "UOO",
    Start_PlaceType=="Off-campus Workplace" & End_PlaceType=="Off-campus Workplace" ~ "UOO",
    Start_PlaceType=="Other" & End_PlaceType=="Off-campus Workplace" ~ "UOO",
    End_PlaceType=="Other" & Start_PlaceType=="Off-campus Workplace" ~ "UOO"), 
    Origin_PA = case_when(
      Start_PlaceType=="Home" ~ "P",
      End_PlaceType!="Home" & Start_PlaceType=="North Carolina State University"~ "P",
      End_PlaceType!="Home" & End_PlaceType!="North Carolina State University"~ "P",
      TRUE ~ "A"),
    Destination_PA = case_when(
      End_PlaceType=="Home" ~ "P",
      Start_PlaceType!="Home"& Start_PlaceType!="North Carolina State University" & End_PlaceType=="North Carolina State University"~ "P",
      TRUE ~ "A"),
    TAZ_P = case_when(
      Origin_PA=="P" ~ ID.x,TRUE ~ ID.y),
    TAZ_A = case_when(
      Origin_PA=="A" ~ ID.x,TRUE ~ ID.y),
    Primary_Mode = as.factor(Mode_1),
    Bicycle = if_else(Primary_Mode == "Bicycle",1,0),
    Bus = if_else((Primary_Mode == "Public Bus / Private Shuttle" | Primary_Mode == "Other, wolfline"),1,0),
    Car = if_else((Primary_Mode == "Driver - Auto / Van / Truck" |  Primary_Mode == "Driver - Auto / Van / Truck, Other" | Primary_Mode == "Passenger - Auto / Van / Truck"),1,0),
    Walk = if_else(Primary_Mode == "Walk", 1,0),
    Other = if_else((Primary_Mode == "Other" | Primary_Mode == "Other, Longboard" | Primary_Mode == "Motorcycle / Motorized Moped or Scooter"),1,0))



```


```{r}

Trip_subset_df<-NCSUtemp_df  %>% 
  filter(Finished==1 & Derived_minus_stated>=-1) %>%
  select(Person_ID,
         On_campus,
         Start_PlaceID,
         Start_PlaceType,
         End_PlaceID,
         End_PlaceType, 
         TAZ_o=ID.x,
         TAZ_d=ID.y,
         Origin_PA,
         Destination_PA,
         TAZ_P,
         TAZ_A,
         Trip_Purpose,
         Purpose,
         Primary_Mode,
         Bicycle,
         Bus,
         Car,
         Walk,
         Other,
         #Distance,
         Weight)


```

#Person-level data
```{r}
Person_subset_df<-NCSU_Person_df %>%
  filter(Finished==1,
         Derived_minus_stated>=-1)%>%
  select(Person_ID,
         On_campus,
         Class_status,
         Employed,
         Have_children,
         Have_car, 
         Have_parking_permit, 
         Not_to_campus_weekdays,
         Stated_trip_rate,
         Derived_trip_rate,
         Derived_minus_stated,
         Graduate,
         Full_time, 
         Weight) %>%
mutate(Job = case_when(Employed=="No" ~ "0",
                    Employed=="Yes, both on and off campus" ~ "3",
                    Employed=="Yes, off campus" ~ "2",
                    Employed=="Yes, on campus" ~ "1"),
       Job_on= case_when(Employed== "Yes, on campus" ~ 1,
                         Employed== "Yes, both on and off campus" ~ 1,
                         TRUE ~ 0),
       Job_off =case_when(Employed== "Yes, off campus" ~ 1,
                         Employed== "Yes, both on and off campus" ~ 1,
                         TRUE ~ 0),
       Permit = case_when(Have_parking_permit=="Yes" ~ 1,
                          TRUE ~ 0))

```

#Socioeconomics
```{r}
# Total BuildingS per university
TotalBuildingS_NCSU <-socioecon %>% 
  filter(!is.na(BuildingS_NCSU)) %>% 
  summarize(sum=sum(BuildingS_NCSU))%>%
  pull(sum)

TotalBuildingS_UNC <-socioecon %>% 
  filter(!is.na(BuildingS_UNC)) %>% 
  summarize(sum=sum(BuildingS_UNC))%>%
  pull(sum)

TotalBuildingS_DUKE <-socioecon %>% 
  filter(!is.na(BuildingS_DUKE)) %>% 
  summarize(sum=sum(BuildingS_DUKE))%>%
  pull(sum)

TotalBuildingS_NCCU <-socioecon %>% 
  filter(!is.na(BuildingS_NCCU)) %>% 
  summarize(sum=sum(BuildingS_NCCU))%>%
  pull(sum)

TotalBuildingS_NCSU
TotalBuildingS_UNC
TotalBuildingS_DUKE
TotalBuildingS_NCCU


# Distribution of BuildingS by TAZ
socioecon2_df <- socioecon %>%
   mutate(Share_Bldg_NCSU=BuildingS_NCSU/TotalBuildingS_NCSU, 
         Share_Bldg_UNC=BuildingS_UNC/TotalBuildingS_UNC,
         Share_Bldg_DUKE=BuildingS_DUKE/TotalBuildingS_DUKE,
         Share_Bldg_NCCU=BuildingS_NCCU/TotalBuildingS_NCCU)

# test -- delete
TAZs_with_buildings<-socioecon2%>%
  mutate(DUKE=if_else(Share_Bldg_DUKE>0,1,0), 
NCSU=if_else(Share_Bldg_NCSU>0,1,0),
NCCU=if_else(Share_Bldg_NCCU>0,1,0),
UNC=if_else(Share_Bldg_UNC>0,1,0))%>%
  filter(!is.na(DUKE),!is.na(NCSU),!is.na(NCCU),!is.na(UNC))%>%
  summarize(NCSU_sum=sum(NCSU),
            NCCU_sum=sum(NCCU),
            UNC_sum=sum(UNC),
            DUKE_sum=sum(DUKE))

TAZs_with_buildings



```

#Datasets for analysis
```{r}
Productions_df<-Trip_subset_df %>%   
  filter(Trip_Purpose!="99" & 
           !is.na(Trip_Purpose))%>% 
  group_by(Person_ID, 
           Trip_Purpose, 
           On_campus,
           #Distance,
           Primary_Mode) %>%
  summarize(Trips=n()) %>%
  left_join(Person_subset_df, by="Person_ID")%>% 
  mutate(Trips_Weighted= Trips * Weight,
         UHC=if_else(Trip_Purpose=="UHC", 1,0),
         UHO=if_else(Trip_Purpose=="UHO", 1,0),
         UCO=if_else(Trip_Purpose=="UCO", 1,0),
         UC1=if_else(Trip_Purpose=="UC1", 1,0),
         UOO=if_else(Trip_Purpose=="UOO", 1,0))%>%
  rename(On_campus=On_campus.x)%>%
  select(Person_ID,
         Trips,
         Trips_Weighted,
         Primary_Mode,
         Trip_Purpose,
         On_campus,
         #Distance,
         Graduate, 
         Full_time,
         Weight,
         Job_on,
         Job_off, 
         UHC,
         UHO,
         UCO,
         UC1,
         UOO)

Productions_forregression_df<-Productions_df%>%
   group_by(Person_ID,
           On_campus, 
           #Distance,
           Graduate, 
           Full_time,
           Job_on,
           Job_off, 
           UHC,
           UHO,
           UCO,
           UC1,
           UOO,
           Weight, 
           Trips)%>%
  summarize(Total_Trips=sum(Trips)) 

Attractions_df<-Trip_subset_df %>% 
  filter(Trip_Purpose!="99" &
           !is.na(Trip_Purpose))%>% 
  group_by(Person_ID, 
           Trip_Purpose, 
           On_campus, 
           TAZ_A, 
           Primary_Mode) %>%
  summarize(Trips=n()) %>%
  left_join(Person_subset_df, by="Person_ID")%>%
  # left_join(socioecon2,by=c("TAZ_A" ="TAZ_NG"))%>%
  mutate(Trips_Weighted= Trips * Weight,
         UHC=if_else(Trip_Purpose=="UHC", 1,0),
         UHO=if_else(Trip_Purpose=="UHO", 1,0),
         UCO=if_else(Trip_Purpose=="UCO", 1,0),
         UC1=if_else(Trip_Purpose=="UC1", 1,0),
         UOO=if_else(Trip_Purpose=="UOO", 1,0))%>%
  rename(On_campus=On_campus.x)


  
#TEMP##################################################
Attractions_byTAZbySegment_df<-Attractions_df %>% 
  ungroup()%>%
  select(TAZ_A,On_campus,Trip_Purpose,Trips,Trips_Weighted)%>%
  filter(Trip_Purpose=="UHO"| Trip_Purpose=="UCO"|Trip_Purpose=="UOO")%>%
  group_by(TAZ_A,
           On_campus,
           Trip_Purpose)%>%
  right_join(.,socioecon2,by=c("TAZ_A" ="TAZ_NG"))%>%
  mutate(Total_W_Trips=sum(Trips_Weighted),
         On_UHOTrips=if_else(Trip_Purpose=="UHO" & On_campus==1,Trips,as.integer(0)),
         On_UCOTrips=if_else(Trip_Purpose=="UCO"& On_campus==1 ,Trips,as.integer(0)),
         On_UOOTrips=if_else(Trip_Purpose=="UOO" & On_campus==1,Trips,as.integer(0)),
         W_On_UHOTrips=if_else(Trip_Purpose=="UHO"& On_campus==1,Trips_Weighted,0),
         W_On_UCOTrips=if_else(Trip_Purpose=="UCO" & On_campus==1,Trips_Weighted,0),
         W_On_UOOTrips=if_else(Trip_Purpose=="UOO" & On_campus==1,Trips_Weighted,0),
         Off_UHOTrips=if_else(Trip_Purpose=="UHO" & On_campus==0,Trips,as.integer(0)),
         Off_UCOTrips=if_else(Trip_Purpose=="UCO"& On_campus==0,Trips,as.integer(0)),
         Off_UOOTrips=if_else(Trip_Purpose=="UOO" & On_campus==0,Trips,as.integer(0)),
         W_Off_UHOTrips=if_else(Trip_Purpose=="UHO"& On_campus==0,Trips_Weighted,0),
         W_Off_UCOTrips=if_else(Trip_Purpose=="UCO" & On_campus==0,Trips_Weighted,0),
         W_Off_UOOTrips=if_else(Trip_Purpose=="UOO" & On_campus==0,Trips_Weighted,0)) 


Attractions_byTAZfinal_df<-Attractions_byTAZbySegment_df%>%
  ungroup()%>%
  select(TAZ_A,
         On_UHOTrips,
         On_UCOTrips,
         On_UOOTrips,
         W_On_UHOTrips,
         W_On_UCOTrips,
         W_On_UOOTrips,
         Off_UHOTrips,
         Off_UCOTrips,
         Off_UOOTrips,
         W_Off_UHOTrips,
         W_Off_UCOTrips,
         W_Off_UOOTrips)%>%
  group_by(TAZ_A)%>%
  summarize_all(sum)%>%
  right_join(.,socioecon2_df,by=c("TAZ_A" ="TAZ_NG"))
  
#summaries

Trips_byTAZ <-Attractions_df%>%
  group_by(TAZ_A)%>%
  summarize(W_Total_Trips=round(sum(Trips_Weighted),digits=2), Trips=round(sum(Trips),digits=2))%>%
  arrange(-W_Total_Trips)


Trips_byTAZ
 
```

#Outputs
```{r}
write_rds(Productions_df,paste0(private_dir, "Productions_df.RDS"))
write_rds(Attractions_df,paste0(private_dir,"Attractions_df.RDS"))
write_rds(Productions_forregression_df,paste0(private_dir,"Productions_forregression_df.RDS"))
write_rds(Attractions_byTAZ_df,paste0(private_dir,"Attractions_byTAZ_df.RDS"))
write_rds(Attractions_byTAZbySegment_df,paste0(private_dir,"Attractions_byTAZbySegment_df.RDS"))
write_rds(Attractions_byTAZfinal_df,paste0(private_dir,"Attractions_byTAZfinal_df.RDS"))
write_rds(Trip_subset_df, paste0(private_dir,"Trip_subset_df.RDS"))
write_rds(Person_subset_df,paste0(private_dir,"Person_subset_df.RDS"))
write_rds(socioecon2_df,paste0(prepared_data_dir,"socioecon2_df.RDS"))
```