---
title: "Auto Ownership"
author: "Caliper Corporation"
date: "June 9, 2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(dplyr.summarise.inform = FALSE)
options(scipen = 999)

library(tidyverse)
library(mltools)
library(data.table)
library(knitr)
library(kableExtra)
```

```{r, include=FALSE}
hh <- read_csv("data/output/_PRIVATE/survey_processing/hh_processed.csv")
```

```{r}
# 1-hot encode variables so they can be alternative specific
create_factors <- hh %>%
  mutate(
    ID = seq(1, n(), 1),
    kid_per_adult = num_children / num_adults,
    num_vehicles = ifelse(num_vehicles >= 4, 4, num_vehicles),
    num_adults = ifelse(num_adults >= 4, 4, num_adults),
    num_workers = ifelse(num_workers >= 3, 3, num_workers),
    num_nwadults = num_adults - num_workers,
    num_children = ifelse(num_children >= 2, 2, num_children),
    num_seniors = ifelse(num_seniors >= 2, 2, num_seniors),
    num_adults = as.factor(num_adults),
    num_workers = as.factor(num_workers),
    num_nwadults = as.factor(num_nwadults),
    num_children = as.factor(num_children),
    num_seniors = as.factor(num_seniors),
    hh_inc_final = as.factor(hh_inc_final)
  ) %>%
  relocate(ID, .before = "hhid") %>%
  as.data.table()
est_tbl <- one_hot(create_factors, dropCols = FALSE)

# Model estimation done in TransCAD. Write the estimation table out to a
# workspace not in the repo
write_csv(
  est_tbl, "../../working_files/auto_ownership/hh_processed.csv",
  na = ""
)
```

## Introduction

Vehicle ownership is a long-term decision that directly impacts daily mode
choice. The choice is influenced by household factors like income and number
of workers, but also by where people choose to live. The auto-ownership model
allows the TRMG2 to be sensitive to these factors and respond to changes in the
future.

## Model Structure

The auto-ownership model in the TRMG2 makes use of variables from the synthetic
population and zonal accessibility to make predictions. The model structure is
a simple multinomial logistic (MNL) regression model with five alternatives:

- 0 Vehicles
- 1 Vehicle
- 2 Vehicles
- 3 Vehicles
- 4+ Vehicles

## Coefficients

Model estimation was performed using TransCADs built-in logit model engine, and
the table below shows the utility terms, coefficients, and goodness of fit.

```{r, include=FALSE}
coeffs <- read_csv("data/input/auto_ownership/ao_final_coefficients.csv") %>%
  mutate(across(.cols = c(veh0:veh4), .fns = ~round(.x, 3))) %>%
  mutate(across(.fns = ~ifelse(is.na(.x), "", .x)))
```

```{r}
coeffs %>%
  kable(digits = 2) %>%
  kable_styling(full_width = FALSE)
```

The coefficients all have the right sign and the relative sizes are intuitive.
One particularly encouraging result of this model is that households with strong
walk and transit accessibility are less likely to own a vehicle and even less
likely to own multiple vehicles. This adds another dimension of model
sensitivity to transit investments. New transit routes will affect long-term
household decisions about auto ownership, which further influence their daily
decisions about mode choice.

Looking in more detail, the trend for coefficients across alternatives is also
intuitive. Large numbers of workers in a household has a small positive impact on the
utility of owning 1 auto, but a large impact on owning 2 or more. High income
households are more likely to own more vehicles. Children make it less likely to
own 3 or more cars, because there are fewer drivers. Finally, the model's Rho^2
shows strong predictive power.