---
title: "Auto Ownership"
author: "Caliper Corporation"
date: "June 19, 2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(dplyr.summarise.inform = FALSE)
options(scipen = 999)

library(tidyverse)
library(mltools)
library(data.table)
library(knitr)
library(kableExtra)
```

```{r, include=FALSE}
hh <- read_csv("data/output/_PRIVATE/survey_processing/hh_processed.csv")
```

```{r}
# 1-hot encode variables so they can be alternative specific
create_factors <- hh %>%
  mutate(
    ID = seq(1, n(), 1),
    kid_per_adult = num_children / num_adults,
    num_vehicles = ifelse(num_vehicles >= 4, 4, num_vehicles),
    num_adults = ifelse(num_adults >= 4, 4, num_adults),
    num_workers = ifelse(num_workers >= 3, 3, num_workers),
    num_nwadults = num_adults - num_workers,
    num_children = ifelse(num_children >= 2, 2, num_children),
    num_seniors = ifelse(num_seniors >= 2, 2, num_seniors),
    num_adults = as.factor(num_adults),
    num_workers = as.factor(num_workers),
    num_nwadults = as.factor(num_nwadults),
    num_children = as.factor(num_children),
    num_seniors = as.factor(num_seniors),
    hh_inc_final = as.factor(hh_inc_final)
  ) %>%
  relocate(ID, .before = "hhid") %>%
  as.data.table()
est_tbl <- one_hot(create_factors, dropCols = FALSE)

# Model estimation done in TransCAD. Write the estimation table out to a
# workspace not in the repo
write_csv(
  est_tbl, "../../working_files/auto_ownership/hh_processed.csv",
  na = ""
)
```

## Introduction

Auto ownership is a long-term decision that directly impacts daily mode
choice. The choice is influenced by household factors like income and number
of workers, but also by where people choose to live. The auto ownership model
allows TRMG2 to be sensitive to these factors and respond to changes in the
future.

Auto ownership is predicted using a discrete choice model. For more details on
general model form, click
[here](https://tfresource.org/topics/Choice_models.html).

## Model Structure

The auto ownership model in TRMG2 makes use of variables from the synthetic
population and zonal accessibility to make predictions. The model structure is
a simple multinomial logistic (MNL) regression model with five alternatives:

- 0 Vehicles
- 1 Vehicle
- 2 Vehicles
- 3 Vehicles
- 4+ Vehicles

## Coefficients

Model estimation was performed using TransCADs built-in logit model engine, and
the table below shows the utility terms, coefficients, and goodness of fit.

```{r, include=FALSE}
coeffs <- read_csv("../master/resident/auto_ownership/ao_coefficients.csv") %>%
  select(Term = Description, Alternative,Coefficient) %>%
  mutate(Coefficient = round(Coefficient, 1)) %>%
  pivot_wider(id_cols = c(Term), names_from = Alternative, values_from = Coefficient) %>%
  mutate(across(.cols = c(v0:v4), .fns = ~round(.x, 3))) %>%
  mutate(across(.fns = ~ifelse(is.na(.x), "", .x)))
```

```{r}
coeffs %>%
  kable(digits = 2) %>%
  kable_styling(full_width = FALSE)
```

Household income categories are defined [here](population_synthesis.html#household-income). Accessibility terms are discussed in detail [here](accessibility.html).

As an example, the 1-vehicle coefficients in the table above translate into
a utility formula that starts like this:

$U = 4.88 + 0.7 * Workers + 0.1 * Seniors - 0.4 * WalkAccess + ...$

The coefficients all have the right sign and the relative sizes are intuitive.
For example, higher-income households are more likely to own more cars. One
particularly encouraging result of this model is that households with strong
walk and transit accessibility are less likely to own a vehicle and even less
likely to own multiple vehicles. This adds another dimension of model
sensitivity to transit investments. New transit routes will affect long-term
household decisions about auto ownership, which further influence their daily
transportation decisions.

Looking in more detail, the trend for coefficients across alternatives is also
intuitive. Large numbers of workers in a household have a small positive impact
on the utility of owning 1 auto, but a large impact on owning 2 or more. High
income households are more likely to own more vehicles. Finally, the model's
Rho^2 of 0.42 shows strong predictive power.

## Calibration

The estimated coefficients produce results that closely matched the survey as
shown in the table below.

```{r, include=FALSE}
# These should come from an application of the model before any calibration
# has been done.
model <- read_csv("data/input/auto_ownership/uncalibrated_auto_results.csv")
```

```{r}
survey <- est_tbl %>%
  group_by(autos = num_vehicles) %>%
  summarize(weight = sum(hh_weight_combined)) %>%
  mutate(pct = weight / sum(weight))

survey %>%
  select(autos, survey = pct) %>%
  left_join(model, by = "autos") %>%
  select(-hh) %>%
  rename(model = pct) %>%
  mutate(
    survey = round(survey * 100, 1),
    model = round(model * 100, 1)
  ) %>%
  rename(Autos = autos, `Survey %` = survey, `Model %` = model) %>%
  kable() %>%
  kable_styling(full_width = FALSE)

```

Caliper created an auto-calibration tool for the auto ownership model and used
it to ensure the model matched the observed shares more closely. Importantly,
the calibration constants required to do this were small as shown below.

```{r, include=FALSE}
calib_coeffs_df <- read_csv("../master/resident/auto_ownership/ao_coefficients.csv")
```

```{r}
calib_coeffs_df %>%
  filter(grepl("Calibration constant", Description)) %>%
  select(Alternative, `Calibration Constant` = Coefficient) %>%
  kable() %>%
  kable_styling(full_width = FALSE)
```


```{r, eval=FALSE}
write_csv(survey, "data/output/auto_ownership/ao_calib_targets.csv")
```

