---
title: "On-board Survey Processing"
author: "Caliper Corporation"
date: "March 15, 2021"
output: 
  html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(dplyr.summarise.inform = FALSE)
options(scipen = 999)

library(tidyverse)
library(knitr)
library(kableExtra)
library(readxl)
library(sf)
```

## Introduction

On-board surveys ask transit riders a series of question while they are on the
bus in order to obtain detailed information about demographics, trip purpose,
modes used, and other variables. These targeted surveys are critical for
understanding transit users in the region. Due to the small size of the transit
market in the Triangle and the small household survey sample size, the household
survey does not contain enough transit trips to estimate the behavior of these
travelers.

In the Triangle, a survey of GoTriangle and DCHC transit was conducted in 2014
(excluding Duke Transit). CAMPO agencies were surveyed in 2015 along with
new GoTriangle routes. These are used to gain insight into the transit market.

## Combining surveys

Unlike the household surveys, the transit surveys were collected by different
companies. Combining the surveys required translating the different field codes
from each into a combined set of values. The surveys collected the same basic
information, which made this process easier. Additionally, because different
routes were collected in each survey, no routes were double counted and no
adjustments to the sample weights were needed.

```{r, include=FALSE}
goraleigh_2015_raw <- read_excel("data/input/_PRIVATE/survey_data/SurveyData_Caliper/2015 GoRaleigh Transit On-Board/1_Data - Original (CDM-Smith)/2015TOB_GoRaleigh_Data_Weighted_Final_2016-04-05_CDMSmith.xlsx")
gotriangle_2014_raw <- read_excel("data/input/_PRIVATE/survey_data/SurveyData_Caliper/2014 GoTriangle Transit On-Board/1_Data - Original/TTA survey data 8-24_reweighted.xlsx")
```

```{r}
# combine surveys
goraleigh_2015 <- goraleigh_2015_raw %>%
  transmute(
    agency = AGENCY,
    route = ROUTE_SURVEYED,
    weight = EXPANSION_FACTOR,
    home_lat = HOME_OR_HOTEL_LAT,
    home_lon = HOME_OR_HOTEL_LON,
    orig_lat = ORIGIN_LAT,
    orig_lon = ORIGIN_LON,
    board_lat = BOARDING_LAT,
    board_lon = BOARDING_LON,
    tod = case_when(
      TIME_ON %in% c("6:00 - 7:00 am", "7:00 - 8:00 am", "8:00 - 9:30 am") ~ "AM",
      TIME_ON %in% c("3:30 - 4:30 pm", "4:30 - 5:30 pm", "5:30 - 6:30 pm") ~ "PM",
      TIME_ON %in% c(
        "9:30 - 10:30 am", "10:30 - 11:30 am", "11:30 am - 12:30 pm", 
        "12:30 - 1:30 pm", "1:30 - 2:30 pm", "2:30 - 3:30 pm"
      ) ~ "MD",
      TRUE ~ "NT"
    ),
    alight_lat = ALIGHTING_LAT,
    alight_lon = ALIGHTING_LON,
    dest_lat = DESTIN_LAT,
    dest_lon = DESTIN_LON,
    access_mode = case_when(
      ACCESS_MODE_CODE == 1 ~ "walk",
      ACCESS_MODE_CODE == 2 ~ "bike",
      ACCESS_MODE_CODE == 4 ~ "sovpnr",
      ACCESS_MODE_CODE == 5 ~ "hovpnr",
      ACCESS_MODE_CODE == 6 ~ "knr",
      TRUE ~ "other"
    ),
    origin_type = case_when(
      ORIGIN_PLACE_TYPE_CODE == 1 ~ "home",
      ORIGIN_PLACE_TYPE_CODE == 2 ~ "work",
      ORIGIN_PLACE_TYPE_CODE == 3 ~ "wr",
      ORIGIN_PLACE_TYPE_CODE == 4 ~ "univ",
      ORIGIN_PLACE_TYPE_CODE == 5 ~ "k12",
      ORIGIN_PLACE_TYPE_CODE == 6 ~ "med",
      ORIGIN_PLACE_TYPE_CODE %in% 8:9 ~ "ome",
      TRUE ~ "other"
    ),
    dest_type = case_when(
      DESTIN_PLACE_TYPE_CODE == 1 ~ "home",
      DESTIN_PLACE_TYPE_CODE == 2 ~ "work",
      DESTIN_PLACE_TYPE_CODE == 3 ~ "wr",
      DESTIN_PLACE_TYPE_CODE == 4 ~ "univ",
      DESTIN_PLACE_TYPE_CODE == 5 ~ "k12",
      DESTIN_PLACE_TYPE_CODE == 6 ~ "med",
      DESTIN_PLACE_TYPE_CODE %in% 8:9 ~ "ome",
      TRUE ~ "other"
    ),
    egress_mode = case_when(
      EGRESS_MODE_CODE == 1 ~ "walk",
      EGRESS_MODE_CODE == 2 ~ "bike",
      EGRESS_MODE_CODE == 4 ~ "sovpnr",
      EGRESS_MODE_CODE == 5 ~ "hovpnr",
      EGRESS_MODE_CODE == 6 ~ "knr",
      TRUE ~ "other"
    ),
    fare_type = case_when(
      FARE_DISCOUNTS_CODE == 1 ~ "full",
      FARE_DISCOUNTS_CODE == 2 ~ "disability",
      FARE_DISCOUNTS_CODE == 3 ~ "youth",
      FARE_DISCOUNTS_CODE == 4 ~ "youth_free",
      FARE_DISCOUNTS_CODE == 5 ~ "senior",
      FARE_DISCOUNTS_CODE == 6 ~ "senior_free",
      FARE_DISCOUNTS_CODE == 7 ~ "pass",
      TRUE ~ "free_route"
    ),
    pass = TRANS_PASS_CARD,
    hh_veh = ifelse(COUNT_VH_HH_CODE == "10P", 10, COUNT_VH_HH_CODE),
    hh_veh = as.numeric(hh_veh),
    hh_size = as.numeric(COUNT_MEMBER_HH_CODE),
    hh_worker = as.numeric(COUNT_EMPLOYED_HH_CODE),
    employed = ifelse(EMPLOYMENT_STATUS_CODE %in% 1:2, "yes", "no"),
    student = case_when(
      EMPLOYMENT_STATUS == 3 ~ "U",
      EMPLOYMENT_STATUS == 7 ~ "K12",
      TRUE ~ "no"
    ),
    license = ifelse(HAS_DRIVE_LICENSE_CODE == 1, "yes", "no"),
    age = case_when(
      AGE_CODE == 1 ~ "child",
      AGE_CODE < 7 ~ "adult",
      TRUE ~ "senior"
    ),
    income = case_when(
      INCOME_CODE <= 5 ~ 1,
      INCOME_CODE <= 9 ~ 2,
      INCOME_CODE <= 11 ~ 3,
      TRUE ~ 4
    ),
    transfers = ifelse(TRANSFERS_FROM_CODE == "4+", 4, as.numeric(TRANSFERS_FROM_CODE))
  )

gotriangle_2014 <- gotriangle_2014_raw %>%
  transmute(
    agency = agency,
    route = route_detail,
    weight = final_weight,
    home_lat = home_lat,
    home_lon = home_lon,
    orig_lat = origin_lat,
    orig_lon = origin_lon,
    board_lat = board_lat,
    board_lon = board_lon,
    tod = case_when(
      board_time %in% c("8 - 9 a.m.", "6 - 7 a.m.", "7 - 8 a.m.") ~ "AM",
      board_time %in% c(
        "1 - 2 p.m.", "3 - 3:30 p.m", "9 - 10 a.m.", "10 - 11 a.m.", 
        "2 - 3 p.m.", "11 - 12 p.m.", "12 - 1 p.m."
      ) ~ "MD",
      board_time %in% c(
        "3:30 - 4 p.m.", "5 - 6 p.m.", "4 - 5 p.m.", "6 - 6:30 p.m."
      ) ~ "PM",
      TRUE ~ "NT"
    ),
    alight_lat = alight_lat,
    alight_lon = alight_lon,
    dest_lat = dest_lat,
    dest_lon = dest_lon,
    access_mode = case_when(
      access == "Walk" ~ "walk",
      access == "Bike" ~ "bike",
      access == "Drove alone and parked" ~ "sovpnr",
      access == "Drove or rode with others and parked" ~ "hovpnr",
      access == "Was dropped off by someone" ~ "knr",
      TRUE ~ "other"
    ),
    origin_type = case_when(
      origin_type == "Home" ~ "home",
      origin_type == "Work" ~ "work",
      origin_type == "College/University" ~ "univ",
      origin_type == "Shopping / Eating / Dining" ~ "ome",
      origin_type == "Other work related place" ~ "wr",
      origin_type == "Medical / Hospital, non work" ~ "med",
      origin_type == "K-12 school" ~ "k12",
      TRUE ~ "other"
    ),
    dest_type = case_when(
      dest_type == "Home" ~ "home",
      dest_type == "Work" ~ "work",
      dest_type == "College/University" ~ "univ",
      dest_type == "Shopping / Eating / Dining" ~ "ome",
      dest_type == "Other work related place" ~ "wr",
      dest_type == "Medical / Hospital, non work" ~ "med",
      dest_type == "K-12 school" ~ "k12",
      TRUE ~ "other"
    ),
    egress_mode = case_when(
      egress == "Walk" ~ "walk",
      egress == "Bike" ~ "bike",
      egress == "Drove alone and parked" ~ "sovpnr",
      egress == "Drove or rode with others and parked" ~ "hovpnr",
      egress == "Was dropped off by someone" ~ "knr",
      TRUE ~ "other"
    ),
    fare_type = case_when(
      fare_type == "Pay the full fare (cash fare or regular pass purchased for yourself)" ~ "full",
      fare_type == "Use a transit pass provided by your employer or university" ~ "pass",
      fare_type == "Use a transfer from another bus" ~ "xfer",
      fare_type == "Pay the Youth Fare" ~ "youth",
      fare_type == "Ride DATA for free with Senior (65+) ID" ~ "senior_free",
      fare_type == "Pay the Triangle Transit Senior Fare" ~ "senior",
      fare_type == "Pay the Person with a Disability Fare" ~ "disability",
      TRUE ~ "free_route"
    ),
    pass = pass,
    hh_veh = hh_veh,
    hh_size = hh_tot,
    hh_worker = hh_emp,
    employed = case_when(
      employment == "Employed full-time" ~ "yes",
      employment == "Employed part-time" ~ "yes",
      TRUE ~ "no"
    ),
    student = case_when(
      student == "Not a student" ~ "no",
      student == "Yes - K-12th grade" ~ "K12",
      TRUE ~ "U"
    ),
    license = license,
    age = case_when(
      age == "18-24" ~ "adult",
      age == "25-34" ~ "adult",
      age == "35-44" ~ "adult",
      age == "45-54" ~ "adult",
      age == "55-64" ~ "adult",
      age == "65+" ~ "senior",
      TRUE ~ "child"
    ),
    # The income breaks in this survey don't match up with the others
    income = case_when(
      income == "Less than $10,000" ~ 1,
      income == "$10,000 - $14,999" ~ 1,
      income == "$15,000 - $19,999" ~ 1,
      income == "$20,000 - $29,999" ~ 1,
      income == "$30,000 - $44,999" ~ 1,
      income == "$45,000 - $59,999" ~ 2,
      income == "$60,000 - $74,999" ~ 2,
      income == "$75,000 - $99,999" ~ 3,
      TRUE ~ 3
    ),
    transfers = transfers
  )

combined <- bind_rows(goraleigh_2015, gotriangle_2014) %>%
  mutate(seqtripid = seq(1, n(), 1)) %>%
  relocate(seqtripid, .before = "agency")
```

## Geocoding

As with the household survey, the transit trip latitude/longitude values were
translated into TRMG2 TAZs. For a transit trip, this includes the home, origin,
destination, boarding, and alighting locations. A sample of this is shown in
the table below:

```{r, include=FALSE}
tazs <- st_read("data/input/tazs/master_tazs.shp") %>%
  st_transform(crs = st_crs('+proj=longlat +datum=WGS84'))
```

```{r}
# create unique list of lat/long points
pts <- tibble(
  lng = c(
    combined$orig_lon,
    combined$dest_lon,
    combined$board_lon,
    combined$alight_lon,
    combined$home_lon
  ),
  lat = c(
    combined$orig_lat,
    combined$dest_lat,
    combined$board_lat,
    combined$alight_lat,
    combined$home_lat
  ),
) %>%
  unite(lng_lat, c(lng, lat)) %>%
  group_by(lng_lat) %>%
  summarize() %>%
  separate(lng_lat, into = c("lng", "lat"), sep = "_") %>%
  filter(lng != "NA") %>%
  mutate(
    lng = as.numeric(lng),
    lat = as.numeric(lat)
  ) %>%
  st_as_sf(
    coords = c("lng", "lat"), 
    crs = st_crs(tazs),
    remove = FALSE
  )
```

```{r, include=FALSE}
geocode_points <- pts %>%
  st_join(tazs %>% select(TAZ = ID)) %>%
  st_drop_geometry()
```

```{r}
add_taz_id <- combined %>%
  left_join(
    geocode_points,
    by = c("home_lon" = "lng", "home_lat" = "lat")
  ) %>%
  rename(home_taz = TAZ) %>%
  left_join(
    geocode_points,
    by = c("orig_lon" = "lng", "orig_lat" = "lat")
  ) %>%
  rename(orig_taz = TAZ) %>%
  left_join(
    geocode_points,
    by = c("board_lon" = "lng", "board_lat" = "lat")
  ) %>%
  rename(board_taz = TAZ) %>%
  left_join(
    geocode_points,
    by = c("alight_lon" = "lng", "alight_lat" = "lat")
  ) %>%
  rename(alight_taz = TAZ) %>%
  left_join(
    geocode_points,
    by = c("dest_lon" = "lng", "dest_lat" = "lat")
  ) %>%
  rename(dest_taz = TAZ)

add_taz_id %>%
  select(route, weight, home_taz:dest_taz) %>%
  mutate(weight = round(weight, 2)) %>%
  head() %>%
  kable() %>%
  kable_styling(full_width = FALSE)
```

```{r}
flip_to_pa <- add_taz_id %>%
  mutate(
    pa_flag = ifelse(dest_type == "home", 0, 1),
    p_taz = ifelse(pa_flag == 1, orig_taz, dest_taz),
    a_taz = ifelse(pa_flag == 0, orig_taz, dest_taz)
  )
```

```{r}
# calculate the transfer rate
xfer_rate <- weighted.mean(flip_to_pa$transfers, flip_to_pa$weight, na.rm = TRUE)
surveyed_trips <- sum(flip_to_pa$weight, na.rm = TRUE)
total_boardings <- surveyed_trips * (1 + xfer_rate)
```

```{r}
add_trip_purp <- flip_to_pa %>%
  mutate(obs_purpose = case_when(
    origin_type == "home" & dest_type == "work" ~ "HBW",
    origin_type == "work" & dest_type == "home" ~ "HBW",
    origin_type == "home" & dest_type == "ome" ~ "HBShop",
    origin_type == "ome" & dest_type == "home" ~ "HBShop",
    origin_type != "home" & dest_type != "home" ~ "NHB",
    TRUE ~ "HBO"
  ))
```

```{r, include=FALSE}
equiv_tbl <- read_csv("data/input/obs_processing/route_name_equiv.csv")
```

```{r}
add_mode <- add_trip_purp %>%
  left_join(equiv_tbl, by = "route") %>%
  # mutate(mode = ifelse(mode == 2, "local", "express"))
  mutate(
    mode = ifelse(mode == 2, "lb", "eb"),
    mode = ifelse(is.na(mode), "eb", mode), # a few GoTriangle routes wind up NA
    access_mode = case_when(
      access_mode == "walk" ~ "w",
      grepl("pnr", access_mode) ~ "pnr",
      grepl("knr", access_mode) ~ "knr",
      TRUE ~ "other"
    ),
    # used for estimation in TC software
    mc_code = case_when(
      mode == "lb" & access_mode == "w" ~ 4,
      mode == "lb" & access_mode == "knr" ~ 5,
      mode == "lb" & access_mode == "pnr" ~ 6,
      mode == "eb" & access_mode == "w" ~ 7,
      mode == "eb" & access_mode == "knr" ~ 8,
      mode == "eb" & access_mode == "pnr" ~ 9
    )
  ) %>%
  relocate(access_mode, .before = mode)
```

## Reasonableness checks {.tabset}

A high level review of the combined survey was performed to make sure processing
had not introduced errors. Some of those checks are shown in the charts below.
All dimensions checked looked reasonable.

### Agency

```{r}
temp <- add_mode %>%
  mutate(agency = ifelse(agency == "TTA", "GoTriangle", agency)) %>%
  group_by(agency) %>%
  summarize(weight = sum(weight)) %>%
  mutate(
    pct = round(weight / sum(weight) * 100, 0)
  )

ggplot(temp, aes(x = reorder(agency, -pct), y = pct)) +
  geom_bar(stat = "identity", aes(fill = agency)) +
  theme(legend.position = "none") + 
  labs(
    title = "Agency Distribution",
    x = "Agency",
    y = "Percent"
  ) +
  theme(plot.title = element_text(hjust = .5))
```


### Time of Day

```{r}
temp <- add_mode %>%
  group_by(tod) %>%
  summarize(weight = sum(weight)) %>%
  mutate(
    pct = round(weight / sum(weight) * 100, 0),
    tod = factor(tod, levels = c("AM", "MD", "PM", "NT"))
  ) %>%
  arrange(tod)

ggplot(temp, aes(x = tod, y = pct)) +
  geom_bar(stat = "identity", aes(fill = tod)) +
  theme(legend.position = "none") + 
  labs(
    title = "Time of Day Distribution",
    x = "Time of Day",
    y = "Percent"
  ) +
  theme(plot.title = element_text(hjust = .5))
```

### Access Mode

```{r}
temp <- add_mode %>%
  group_by(access_mode) %>%
  summarize(weight = sum(weight)) %>%
  mutate(
    pct = round(weight / sum(weight) * 100, 0),
  )

ggplot(temp, aes(x = reorder(access_mode, -pct), y = pct)) +
  geom_bar(stat = "identity", fill = "gold") +
  labs(
    title = "Access Mode Distribution",
    x = "Access Mode",
    y = "Percent"
  ) +
  theme(plot.title = element_text(hjust = .5))
```

### Fare Type

```{r}
temp <- add_mode %>%
  group_by(fare_type) %>%
  summarize(weight = sum(weight)) %>%
  mutate(
    pct = round(weight / sum(weight) * 100, 0)
  )

ggplot(temp, aes(x = reorder(fare_type, -pct), y = pct)) +
  geom_bar(stat = "identity", aes(fill = fare_type)) +
  theme(legend.position = "none") + 
  labs(
    title = "Fare Type Distribution",
    x = "Fare Type",
    y = "Percent"
  ) +
  theme(plot.title = element_text(hjust = .5))
```


## Survey Assignment

For an additional check on both the on-board survey processing and the model's
route network, the on-board survey was assigned. This involved processing the
survey into the appropriate matrices by time of day, transit mode, and accesss
mode. The model assignment macros in TransCAD were used to assign these to the
route system, and the results were compared to reported ridership by route.

```{r, eval=FALSE}
# This code chunk was used to create transit matrices for a test assignment.
# It does not need to be run each time the page is knit, but is kept as
# a reference.
library(caliperR)
source("R/create_obs_matrices.R")
temp <- add_mode %>%
  filter(access_mode != "other", !is.na(mode)) %>%
  group_by(board_time, access_mode, mode) %>%
  nest()

create_obs_matrices(temp)
```

```{r, include=FALSE}
model_boardings <- read_csv(
  "data/input/obs_processing/model_boardings_and_alightings.csv") %>%
  filter(tod == "Daily") %>%
  select(rts_name, model_boardings = On)

reported_boardings <- read_excel(
  "data/input/obs_processing/reported_ridership.xlsx", sheet = "reported")
reported_equiv <- read_excel(
  "data/input/obs_processing/reported_ridership.xlsx", sheet = "equiv")
```

```{r}
comparison <- reported_equiv %>%
  left_join(reported_boardings, by = c("reported_route" = "route")) %>%
  group_by(reported_route) %>%
  mutate(
    count = n(),
    ridership_2016 = ridership_2016 / count
  ) %>%
  ungroup() %>%
  select(model_route, reported = ridership_2016) %>%
  left_join(model_boardings, by = c("model_route" = "rts_name")) %>%
  filter(!is.na(model_boardings), !is.na(reported)) %>%
  mutate(
    diff = model_boardings - reported,
    abs_diff = abs(diff),
    pct_diff = round(diff / reported * 100, 2)
  ) %>%
  # Duke routes are not included in either survey combined, so don't include them
  # for RMSE calculation
  filter(!grepl("DUKE", model_route))

route_prmse <-  round(
  Metrics::rmse(comparison$model_boardings, comparison$reported) / 
    mean(comparison$reported) * 100, 2
)

agency_comp <- comparison %>%
  separate(
    model_route, into = c("agency", "route"), sep = " ", remove = FALSE,
    extra = "merge"
  ) %>%
  group_by(agency) %>%
  summarize(reported = sum(reported), model_boardings = sum(model_boardings)) %>%
  mutate(
    reported = round(reported, 0),
    model_boardings = round(model_boardings, 0),
    diff = model_boardings - reported,
    abs_diff = abs(diff),
    pct_diff = round(diff / reported * 100, 2)
  )

agency_prmse <- round(
  Metrics::rmse(agency_comp$model_boardings, agency_comp$reported) / 
    mean(agency_comp$reported) * 100, 2
)
```

The percent root mean square error (%RMSE) for route-level boardings was
`r route_prmse`%. The agency-level %RMSE was `r round(agency_prmse, 2)`%. The
table below shows the agency-level comparison.

```{r}
agency_comp %>%
  filter(!(agency %in% c("CLX", "FRX"))) %>%
  rename(
    Agency = agency,
    Observed = reported,
    Modelled = model_boardings,
    Difference = diff,
    `Absolute Difference` = abs_diff,
    `Percent Difference` = pct_diff
  ) %>%
  kable(format.args = list(big.mark = ",")) %>%
  kable_styling(full_width = FALSE)
```

These differences are large, but manual path checking did not uncover any
errors. This test assignment was performed on a network with a rough initial
guess of travel times, which likely contributed to these differences. More
attention will be paid to agency- and route-level ridership comparisons during
model validation. It may be that fares or other route attributes received from
stakeholders need to be adjusted. One positive result of this early test was
that all routes received ridership, which indicates no critical accessibility
flaws in the network.

*TODO: Revise this section after validation*

```{r, eval=FALSE}
write_csv(add_mode, "data/output/_PRIVATE/obs_processing/obs_processed.csv", na = "")
write_csv(comparison, "data/output/obs_processing/ridership_comparison.csv")
```

