---
title: "Non-Motorized Choice"
author: "Caliper Corporation"
date: "August 9, 2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(dplyr.summarise.inform = FALSE)
options(scipen = 999)

library(tidyverse)
library(knitr)
library(kableExtra)
```

## Intro

TRMG2 uses a disaggregate, binary choice model to split trips into
motorized and non-motorized categories. It is common to see the non-motorized
mode as a nest within the larger mode choice model, but it was separated for
two primary reasons:

1. To use a disaggregate model to predict walk trips.
2. To simplify the mode choice model and improve prediction power.

In TRMG2, the non-motorized model takes place after trip production but
before trips are aggregated for destination and motorized mode choice models. The
advantage with a disaggregate model is that person and household level
variables were available to improve non-motorized predictions. A potential
downside of this approach is that the prediction does not know the mileage or travel time of the trip. This downside is mitigated by using accessibility measures, which in this
case are logsums of the gravity model used to distribute walk trips. In this
way, aggregate information about trip lengths is included. Further, a person
living downtown will make more walk trips than one living in rural areas (all
else equal).

The second benefit of a separate non-motorized model is a simplification of
the larger mode choice nesting structure. For more detail on that model, and for
information on logit choice models in general, see the resident mode choice
page.

## Estimation Results

All the models that follow are binary choice models with motorized as the
reference alternative. Negative coefficients mean that non-motorized travel
is less likely. Positive coefficients mean that non-motorized travel is more
likely.

Most coefficients are self-explanatory, but two need differentiation:

- Walk Accessibility: this is a measure of how accessible other zones are
  to the home zone (see the [accessibility](accessibility.html) page).
- Walkability: this is a measure of how walkable the home zone itself is. It
  encapsulates information on mixed use, the density of the street grid, and
  other factors. See the accessibility documentation for more detail.

### Work tours

The W_HB_W and W_HB_O trip types were estimated in a single model. Behaviorally,
the decision to walk to work is made at the tour level and affects all trips
made along that tour. A model was not estimated for W_HB_WR trips. Instead,
these trips for deliveries, lawn care, and other activities are all assumed to
be motorized by the model. Similarly, all K12 escort trips on work tours
(W_HB_EK12) are motorized as well.

Children in the household or being a senior make it less likely that trips on a 
work tour will be non-motorized. Higher car ownership has the same affect.
Walk accessibility has a positive relationship, meaning that people who live
in urban centers with high walk accessibility are more likely to make walk
trips. All coefficients make behavioral sense and the model fit is strong.

```{r}
tribble(
  ~Term, ~Coefficient, ~Tstat,
  "Is Senior", -.595, -1.84,
  "Children in HH", -.641, -4.59,
  "Vehicles per Adult", -1.44, -8.71,
  "Walk Accessibility", 1, 11.3,
  "Constant", -2.12, -11.73
) %>%
  kable() %>%
  kable_styling(full_width = FALSE)
```

Rho-squared: 0.82

### Non-work tours

#### N_HB_K12

The decision to walk to school is driven by vehicle ownership, walk
accessibility, and whether or not the child is old enough to drive herself to
school. A high rho-squared metric shows a strong model fit.

```{r}
tribble(
  ~Term, ~Coefficient, ~Tstat,
  "Vehicles per Adult", -.767, -2.53,
  "Walk Accessibility", .585, 2.9,
  "Age 16 to 18", -.359, -1.1,
  "Constant", -2.71, -8.08
) %>%
  kable() %>%
  kable_styling(full_width = FALSE)
```

Rho-squared: 0.80

#### N_HB_OME

The choice for shopping, dining, and maintenance trips is driven by similar
variables as previous trips, but also new ones including income and walkability.
Income has a positive impact on walking, which may at first appear counter-
intuitive. However, many of the factors like vehicle ownership, work status,
and accessibility that are often related to income are already controlled for.

```{r}
tribble(
  ~Term, ~Coefficient, ~Tstat,
  "Vehicles per Adult", -1.94, -13.6,
  "Walk Accessibility", .547, 5.12,
  "Is Worker", .311, 3.03,
  "Income per Person", 0.000008, 4.48,
  "Walkability", 7.3, 3.28,
  "Constant", -2.58, -9.32
) %>%
  kable() %>%
  kable_styling(full_width = FALSE)
```

Rho-squared: 0.72

#### N_HB_OD_Long

These discretionary trips with long activity duration are influenced by many
factors. Children are less likely to walk or bike, higher accessibility leads
to more walk trips, and the other relationships make sense.

```{r}
tribble(
  ~Term, ~Coefficient, ~Tstat,
  "Vehicles per Adult", -1.13, -8.81,
  "Walk Accessibility", .417, 4.54,
  "Is Child", -.465, -3.9,
  "Is Worker", -.283, -3.02,
  "Income per Person", 0.0000044, 2.76,
  "Walkability", 3.7, 2.19,
  "Constant", -2.58, -9.32
) %>%
  kable() %>%
  kable_styling(full_width = FALSE)
```

Rho-squared: 0.59

#### N_HB_OD_Short

This is the only model that struggles to predict non-motorized behavior as shown
by the low rho-squared value. These are "quick stop" discretionary trips, and the
poor fit means that choosing to walk for these trips is based on factors
not collected in the survey. While the t-stat for walkability was not large,
the coefficient makes sense behaviorally and provides sensitivity to land use.

```{r}
tribble(
  ~Term, ~Coefficient, ~Tstat,
  "Is Senior", -.361, -4.3,
  "Is Worker", -.188, -3.13,
  "Walkability", .518, .36,
  "Children in HH", -.579, -8.48,
  "No Autos", .89, 4.16,
  "Income per Person", 0.00001, 8.41,
  "Constant", -.22, -1.17
) %>%
  kable() %>%
  kable_styling(full_width = FALSE)
```

Rho-squared: 0.04

#### N_HB_OMED

The model for medical trips is simple, but does provide important sensitivity
for zero-vehicle households and walk accessibility. It also effectively
prohibits seniors from walking. The rho-squared is large, but this is inflated
given that over 99% of medical trips are motorized. (Randomly assigning 1 out
of every 100 medical trips to walk would also have a high rho-squared value.)

```{r}
tribble(
  ~Term, ~Coefficient, ~Tstat,
  "No Autos", 1.73, .9,
  "Is Senior", -157479, -157479,
  "Walk Accessibility", 4.64, 2.19,
  "Constant", -11.2, -2.42
) %>%
  kable() %>%
  kable_styling(full_width = FALSE)
```

Rho-squared: 0.98

## Calibration

Caliper calibrated the model constants to ensure that the final share
of non-motorized trips matched the survey. This was done using a auto-calibration
utility included in the model's drop down menu. Caliper then reviewed the results
to check that the calibration constants were reasonable and did not dominate
model sensitivity.

The parameter file for N_HB_K12_All trips is shown below as an example. Note that
the estimated constant is preserved separately from the calibration constant.
The additional constant from calibration is small, which means the model has not
sacrificed sensitivity to achieve accuracy in the base year.

```{r}
param_df <- read_csv("../master/resident/nonmotorized/n_hb_k12_all.csv")

param_df %>%
  kable() %>%
  kable_styling(full_width = FALSE)
```

The non-motorized shares resulting from the model are shown below. Each matches
the survey shares.

```{r}
final_shares <- read_csv("data/input/nm/model_nm_summary.csv") %>%
  select(trip_type, nm_share)

colnames(final_shares) <- c("Trip Type", "NM Share (%)")

final_shares %>%
  kable() %>%
  kable_styling(full_width = FALSE)
```

