---
title: "Survey Processing"
author: "Caliper Corporation"
date: "11/5/2020"
output: 
  html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(dplyr.summarise.inform = FALSE)
options(scipen = 999)

library(tidyverse)
library(broom)
library(readxl)
library(ggplot2)
library(plotly)
library(leaflet)
library(sf)
library(geosphere)
library(knitr)
library(scales)
library(kableExtra)
```

```{r read_data, message=FALSE}
# Household files
hh_2016_raw <- read_excel("data/input/_PRIVATE/survey_data/SurveyData_Caliper/2016 Recurrent HTS/1_Data - Original/2016HTS_Data_Main_Households_Final_2016-06-22_RSG.xlsx")
# I exported an xlsx from the access database these were in
combined_hh_weights_2016 <- read_excel("data/input/_PRIVATE/survey_data/SurveyData_Caliper/2016 Recurrent HTS/1_Data - Original/combined_hh_weights_2016.xlsx")
hh_2016_raw <- hh_2016_raw %>%
  left_join(combined_hh_weights_2016, by = c("hhID" = "hhid"))
hh_2018_raw <- read_excel("data/input/_PRIVATE/survey_data/SurveyData_Caliper/2018 Recurrent HTS/1_Data - Original/2018RHTS_HH_2020-04-28.xlsx")

# Person files
per_2016_raw <- read_excel("data/input/_PRIVATE/survey_data/SurveyData_Caliper/2016 Recurrent HTS/1_Data - Original/2016HTS_Data_Main_Persons_Final_2016-07-01_RSG.xlsx")
per_2018_raw <- read_excel("data/input/_PRIVATE/survey_data/SurveyData_Caliper/2018 Recurrent HTS/1_Data - Original/2018RHTS_Person_2020-04-28.xlsx")

# Trip files
trips_2016_raw <- read_excel("data/input/_PRIVATE/survey_data/SurveyData_Caliper/2016 Recurrent HTS/1_Data - Original/2016HTS_Data_Main_Trips_Final_2016-06-22_RSG.xlsx")
# I exported an xlsx from the access database these were in
combined_trip_weights_2016 <- read_excel("data/input/_PRIVATE/survey_data/SurveyData_Caliper/2016 Recurrent HTS/1_Data - Original/combined_trip_weights_2016.xlsx") %>%
  mutate(
    hhid = as.numeric(hhid)
    # personnum = str_pad(personnum, "2", "left", "0"),
    # tripnum = str_pad(tripnum, "3", "left", "0"),
    # hhID = paste0(hhid, personnum, tripnum)
  )
trips_2016_raw <- trips_2016_raw %>%
  left_join(
    combined_trip_weights_2016, 
    by = c("hhID" = "hhid", "personNum" = "personnum", "tripNum" = "tripnum")
  )
trips_2018_raw <- read_excel("data/input/_PRIVATE/survey_data/SurveyData_Caliper/2018 Recurrent HTS/1_Data - Original/2018RHTS_Trip_2020-04-28.xlsx")
```

## Introduction

The Institute for Transportation Research and Education (ITRE) on behalf of the
Triangle Region stakeholders, began conducting rolling household travel surveys
in 2016. For the TRMG2 estimation, both the 2016 and 2018 surveys are available.
This page documents the combination of those surveys along with any general
processing steps not specific to a particular model.

## Survey combination

When RSG delivered the 2018 survey, they provided weighting and expansion factors
for 2018 using two data sets:

  * Only the 2018 samples
  * A combination of 2016 and 2018 samples
  
The combined data set expansion was controlled using 2018 American Community
Survey (ACS) marginals. For TRMG2, with a 2016 base year, these weights were
close enough. For example, the table below shows that the original and combined
weights for the 2016 records produce effectively the same distribution of
household size.

```{r}
remove_na_2016 <- hh_2016_raw %>%
  filter(!is.na(hhweight))
remove_na_2018 <- hh_2018_raw %>%
  filter(!is.na(hh_weight_2018))

remove_na_2016 %>%
  mutate(hhsize = ifelse(hhsize > 5, 5, hhsize)) %>%
  group_by(hhsize) %>%
  summarize(
    weight_2016 = sum(hhweight),
    weight_combined = sum(hh_weight_combined)
  ) %>%
  mutate(
    pct_16 = round(weight_2016 / sum(weight_2016), 3),
    pct_16 = percent(pct_16),
    pct_combined = round(weight_combined / sum(weight_combined), 3),
    pct_combined = percent(pct_combined)
  ) %>%
  select(
    `HH Size` = hhsize, `2016 weights` = pct_16,
    `Combined Weights` = pct_combined
  ) %>%
  kable() %>%
  kable_styling(full_width = FALSE)
```

For models where absolute counts of households matter (like trip production), the
weights will be scaled down to match the 2016 socio-economic (SE) data's total
household count.

### Household table

The 2016 survey's household table did have slight differences in the field names
and definitions from the 2018 survey. Compared to the person and trip tables,
the household table was easy to combine and did not result in losing any
information.

```{r combine_hhs}
prep_2016 <- remove_na_2016 %>%
  rename(hhid = hhID) %>%
  mutate(
    year = 2016,
    hh_income_broad = ifelse(
      hh_income_broad %in% c(5, 99), hhincome_imputed, hh_income_broad
    )
  ) %>%
  select(-hhweight, -hhincome_imputed, -secondhome)
prep_2018 <- remove_na_2018 %>%
  mutate(
    year = 2018,
    home_loc_puma = str_pad(home_loc_puma, 5, "left", pad = "0"),
    home_loc_tract = str_pad(home_loc_tract, 6, "left", pad = "0"),
    hh_income_broad = ifelse(
      imputed_income_flag == 1 | hh_income_broad == 99,
      hh_income_imputed, hh_income_broad
    )
  ) %>%
  select(
    -imputed_income_flag, -hh_weight_2018, -hh_income_imputed,
    -personhh_income_flag
  )

hh_combined <- bind_rows(
  prep_2016, prep_2018
)
```

### Person table

The person table changed substantially between 2016 and 2018. As an example,
race included 6 categories in 2016 and was contained in a single column. In
2018, race only had 5 categories and was contained in 5 separate binary columns
(e.g. `race_white` = 1 or 0). These formats had to be rectified, and the "Two+"
race category from 2016 collapsed into "Other", before the tables could be
combined.

In addition, some fields in the 2016 table were not present in 2018 (and vice
versa). The `num_transitpass` field kept track of how many transit passes the
person used per week in 2016. This field was dropped in 2018. While several
fields were dropped, none were critical to model estimation.

```{r combine_persons}
# combine persons

# # Check for field discrepancies
# bind_rows(
#   per_2016_raw %>% mutate(year = 2016),
#   per_2018_raw %>% mutate(year = 2018)
# ) %>%
#   group_by(year) %>%
#   summarize_all(~sum(is.na(.))) %>%
#   View()

prep_per_2016 <- per_2016_raw %>%
  mutate(
    race = case_when(
      race == 1 ~ "Asian",
      race == 2 ~ "Black",
      race == 3 ~ "White",
      race == 4 ~ "Other",
      race == 5 ~ "Other",
      race == 6 ~ "Noanswer"
    )
  ) %>%
  rename(
    hhid = hhID,
    personid = personID,
    personnum = personNum,
    worker_earnings = worker_income,
    numtrips = num_trips
  ) %>%
  # remove fields that weren't collected in 2018
  select(-c(
    num_transitpass:work_transit, walk_freq,
    hhweight, person_weight_final,
    smartphone_iphone, smartphone_android, android_age
  )) %>%
  mutate(year = 2016)

prep_per_2018 <- per_2018_raw %>%
  select(-c(
    segment:home_loc_region,
    school_taz, school_region,
    work_taz_1, work_region_1, work_taz_2, work_region_2,
    secondhome_taz, secondhome_region,
    hh_income_cat, hh_weight_combined, diary_order, copied_trips_available,
    copied_trips_confirmed, user_ismobiledevice, education_flag,
    school_mode_flag, hh_weight_2018
  )) %>%
  # handle other fields
  mutate(
    personid = as.character(personid),
    year = 2018
  )

# collapse race fields into a single column
race_collapse <- prep_per_2018 %>%
  select(personid, race_asian:race_noanswer) %>%
  mutate(race_noanswer = ifelse(is.na(race_noanswer), 1, race_noanswer)) %>%
  pivot_longer(
    cols = race_asian:race_noanswer,
    names_to = "race"
  ) %>%
  filter(!is.na(value)) %>%
  group_by(personid) %>%
  # Collapse 2+ races into "other"
  mutate(
    race_count = sum(value),
    value = ifelse(race_count > 1, 0, value),
    value = ifelse(race_count > 1 & race == "race_other", 1, value)
  ) %>%
  filter(value == 1) %>%
  mutate(
    race = gsub("race_", "", race),
    race = str_to_title(race)
  ) %>%
  select(-value, -race_count)

prep_per_2018 <- prep_per_2018 %>%
  left_join(race_collapse, by = "personid") %>%
  select(-c(race_asian:race_noanswer))

per_combined <- bind_rows(
  prep_per_2016,
  prep_per_2018
)  
```

### Trip table

Similar to the person table, the trip tables from 2016 and 2018 contained
different fields and field definitions, which complicated the combination
process. Importantly, the definition of activity purpose changed slightly
between the two years. Both columns were left in the combined table, to be
processed in the next step.

```{r combine_trips}
prep_trip_2016 <- trips_2016_raw %>%
  mutate(
    year = 2016,
    personNum = str_pad(personNum, 2, "left", "0"),
    hhID = as.character(hhID),
    personid = paste0(hhID, personNum),
    tripNum = str_pad(tripNum, 3, "left", "0"),
    tripid = paste0(hhID, personid, tripNum)
  ) %>%
  rename(
    hhid = hhID, personnum = personNum, tripnum = tripNum,
    o_activity_2016 = o_activity, d_activity_2016 = d_activity,
    vehiclenum = vehicleNum
  ) %>%
  select(-c(
    tripweight1:tripweight4,
    transit_access:transit_passtype, 
    o_activity_derived, d_activity_derived, tollroad_cost, 
    tollroad_cost_dontknow, trip_flag, gpsfact_age, gpsfact_inc, gpsfact_purp
  ))

prep_trip_2018 <- trips_2018_raw %>%
  mutate(
    year = 2018,
    hhid = as.character(hhid),
    personnum = as.character(personnum),
    personid = str_pad(personid, 2, "left", "0"),
    tripnum = str_pad(tripnum, 3, "left", "0"),
    tripid = as.character(tripid)
  ) %>%
  select(-c(
    last_trip, segment, origin_taz, origin_campus, origin_region,
    destination_taz, destination_campus, destination_region,
    departure_string, arrival_string, o_activity_itre, o_activity_final,
    d_activity_itre, d_activity_final,
    transit_access:transit_passtype, 
    quality_flag, trip_weight_2018
  )) %>%
  rename(o_activity_2018 = o_activity, d_activity_2018 = d_activity)

trip_combined <- bind_rows(
  prep_trip_2016,
  prep_trip_2018
) %>%
  mutate(
    hhid = as.numeric(hhid),
    personnum = as.numeric(personnum),
    tripnum = as.numeric(tripnum)
  ) %>%
  select(hhid, personid, tripid, personnum, tripnum, everything()) %>%
  left_join(
    per_combined %>% select(personid, age),
    by = "personid"
  ) %>%
  mutate(party_size = num_hh_members + num_nonhh)
```

## Basic checks

A common task in household survey processing is to remove households for various
reasons. For instance, travel models estimate weekday travel, so households
surveyed on the weekend must be removed. Households may have incomplete travel
diaries, be located outside the model region, or have other issues that means
they must be removed from estimation set.

### Day of week

Thanks to careful survey design, none of the Triangle households were surveyed
on the weekend. The chart below shows the days surveyed (1 is Monday).

```{r}
temp <- hh_combined %>%
  group_by(year, travelday) %>%
  summarize(samples = n())

ggplot(temp, aes(x = travelday, y = samples, fill = as.factor(travelday))) +
  geom_bar(stat = "identity") +
  guides(fill = FALSE) +
  facet_wrap(~year)
```

Starting in 2018, the survey was improved to only collect travel diaries in the
middle of the week. This improvement recognizes that Monday and Friday are often
different from midweek, which can be due to holidays, vacations, alternative
work schedules, or other reasons. This is a general rule of thumb, however, and
is not necessarily true of all places at all times.

A quick t-test was done to see if travel on Monday and Friday was statistically
different from the mid-week (Tuesday through Thursday). The result suggests that
trip making is not different on Monday and Friday, which means we can keep all
surveyed households in our estimation set.

```{r}
# Compare Friday to not Friday
t.test(
  hh_combined$num_trips[hh_combined$travelday %in% c(1,5)],
  hh_combined$num_trips[!(hh_combined$travelday %in% c(1,5))],
  var.equal = TRUE # if false then performs Welch's t-test
) %>%
  tidy() %>%
  # rename(`mean T-Th` = estimate1, `M & F` = estimate2) %>%
  select(estimate1:p.value) %>%
  pivot_longer(
    cols = estimate1:p.value, names_to = "Parameter", values_to = "Value"
  ) %>%
  mutate(
    Parameter = c(
      "Average Trips/HH (Mon/Fri)",
      "Average Trips/HH (Midweek)",
      "t-stat",
      "p.value"
    )
  ) %>%
  kable(digits = 2) %>%
  kable_styling(full_width = FALSE)
```

### Household locations

It is also important to make sure that only households in your model region are
used for estimation. In addition to the obvious reasons, it will also not be
possible to collect a model-based skim of impedance for these households. The
map below shows the locations of all households included in the survey. Each
falls within the model boundary.

```{r}
shp <- hh_combined %>%
  select(home_loc_lat, home_loc_lng) %>%
  st_as_sf(
    coords = c("home_loc_lng", "home_loc_lat"), crs = 4326, agr = "constant"
  )
boundary_shp <- st_read(
  "data/input/model_boundary/boundary 2020-11-02.shp",
  quiet = TRUE
)
boundary_shp <- st_transform(boundary_shp, st_crs(shp))
bbox <- as.list(st_bbox(boundary_shp))
lng <- (bbox$xmin - bbox$xmax) / 2 + bbox$xmax
lat <- (bbox$ymax - bbox$ymin) / 2 + bbox$ymin

leaflet(
  options = leafletOptions(zoomControl = FALSE, minZoom = 9, maxZoom = 9)
) %>%
  addTiles() %>%
  addPolygons(
    data = boundary_shp,
    color = "#444444", weight = 1, smoothFactor = 0.5
  ) %>%
  addCircleMarkers(
    data = shp, radius = 1, weight = 3,
    opacity = .2, fillOpacity = .2
  ) %>%
  setView(lng, lat, zoom = 9)
```

### Weight distribution

It is important to check the weighting and expansion of a survey for extreme
weight values. To do this, the weights must first be normalized based on the
expected average weight per sample. A simple comparison of two hypothetical
surveys explains why:

1. A survey with 100 samples of a population of 100,000: Average weight = 1,000
2. A survey with 1000 samples of a population of 100,000: Average weight = 100

A household weight of 1,000 in the first survey is reasonable based on the
number of samples, but it is 10 times the expected weight in the second survey.
For the 2016+2018 combined survey, the average weight is shown below.

```{r}
temp <- hh_combined %>%
  summarize(
    Samples = n(),
    `Total Households` = sum(hh_weight_combined),
    `Average Weight` = `Total Households` / Samples
  ) %>%
  pivot_longer(cols = everything())

weight_mean <- mean(hh_combined$hh_weight_combined)
colnames(temp) <- c("", "")

temp %>%
  kable(digits = 0) %>%
  kable_styling(full_width = FALSE)
```

The first histogram below shows the distribution of weights divided by the mean
weight. While the largest weight is 15x the mean weight, this is an outlier and
most are within 6x. This is a reasonable result and indicates and well-designed
sample plan that collected adequate numbers of households in all the demographic
groups of interest.

```{r}
ggplot(hh_combined, aes(x=hh_weight_combined / weight_mean)) + 
  geom_histogram(color="black", fill="orange", bins = 10) +
  xlab("Weight / Average Weight")
```

The second histogram below inverts the factor above and divides the average
weight by the individual households weights. Large factors in this histogram
would indicate the presence of small weights. The smallest weight (17.3) is 14%
of the average weight, suggesting a modest over sample, but is well within
reason.

```{r}
ggplot(hh_combined, aes(x= weight_mean/ hh_weight_combined)) + 
  geom_histogram(color="black", fill="orange", bins = 10) +
  xlab("Average Weight / Weight")
```

After a review of RSGs expansion procedure and the review of weights above,
Caliper determined that the weights were appropriate for use in model
estimation.

## Trip processing

### Activity purpose

The atomic model purposes are shown below. These represent the most fundamental
purposes and may be aggregated during the estimation of specific model
components.

- H: Home (non-work activities at home)
- W: Work
- WR: Work-related (e.g. making a delivery)
- SCH: School. Clarified explicitly in the person table into:
  - PreK: Pre-Kindergarten
  - K12: Kindergarten through 12th grade
  - U: University or College
- SHP: Shop
- EAT: Dining out
- PU: Pick up
- DO: Drop off
- X: Transfer modes (e.g. at a bus stop)
- O: Other

The table below shows how the survey purposes (for both years) were collapsed
into the purposes above.

```{r, include=FALSE}
activity_2016 <- read_csv("data/input/survey_processing/activity_2016.csv")
activity_2018 <- read_csv("data/input/survey_processing/activity_2018.csv")
```

```{r}
activity_2016 %>%
  left_join(activity_2018, by = "code") %>%
  rename(
    `Activity Code` = code,
    `2016 Description` = activity_long.x,
    `2016 Equivalence` = activity.x,
    `2018 Description` = activity_long.y,
    `2018 Eqiuvalence` = activity.y
  ) %>%
  kable() %>%
  kable_styling(bootstrap_options = "striped")
```

```{r activity_purpose}
act_to_join_2016 <- activity_2016 %>% select(-activity_long)
act_to_join_2018 <- activity_2018 %>% select(-activity_long)

convert_purpose <- trip_combined %>%
  left_join(act_to_join_2016, by = c("o_activity_2016" = "code")) %>%
  rename(o_activity1 = activity) %>%
  left_join(act_to_join_2016, by = c("d_activity_2016" = "code")) %>%
  rename(d_activity1 = activity) %>%
  left_join(act_to_join_2018, by = c("o_activity_2018" = "code")) %>%
  rename(o_activity2 = activity) %>%
  left_join(act_to_join_2018, by = c("d_activity_2018" = "code")) %>%
  rename(d_activity2 = activity) %>%
  mutate(
    o_activity = ifelse(is.na(o_activity1), o_activity2, o_activity1),
    d_activity = ifelse(is.na(d_activity1), d_activity2, d_activity1),
  ) %>%
  select(-c(o_activity1:d_activity2)) %>%
  mutate(
    mode = case_when(
      mode == 1 ~ "auto_own",
      mode == 2 & mode_othauto == 1 ~ "auto_work",
      mode == 2 & mode_othauto == 2 ~ "auto_rental",
      mode == 2 & mode_othauto == 3 ~ "auto_taxi",
      mode == 2 & mode_othauto == 7 ~ "auto_tnc",
      mode == 2 ~ "auto_other",
      mode == 3 ~ "bus",
      mode == 4 ~ "walk",
      mode == 5 ~ "bike",
      mode == 6 ~ "other"
    )
  ) %>%
  left_join(
    per_combined %>% select(personid, school),
    by = "personid"
  ) %>%
  mutate(
    # some people marked 1-yo old as student with school at home. Not what
    # we are interested in.
    # student_type = ifelse(is.na(school), 0, 1),
    student_type = case_when(
      school >= 4 & school <= 9 ~ "K12",
      school >= 10 & school <= 13 ~ "U",
      !is.na(school) ~ "PreK",
      TRUE ~ ""
    ),
    o_activity = case_when(
      o_activity != "SCH" ~ o_activity,
      student_type != "" ~ student_type,
      TRUE ~ "SCH?"
    ),
    d_activity = case_when(
      d_activity != "SCH" ~ d_activity,
      student_type != "" ~ student_type,
      TRUE ~ "SCH?"
    ),
    homeschooled_flag = ifelse(school %in% c(5, 7, 9), 1, 0)
  )
```


### Places

As explained above, the definition of a work activity in the 2016 survey did not
specify whether work was performed in or outside the home. Caliper computed a
place code for every trip end independent of activity.

-H: Home
-W: Work
-S: School
-O: Other

Home and school locations were derived by comparing the activity location
(long/lat) with the home and school locations. Trip ends within 100 meters of
the home were tagged to the appropriate place type. This approach gave a more
accurate result than depending on the names each person gave their origin and
destination, which varied widely. It also handled several cases where
respondents separated places like "front yard" and "shed" from their physical
home.

Work locations could not be handled the same way. Not only do people have
multiple jobs, but many had jobs without fixed work locations (e.g. nanny). For
this reason, we relied on the stated activity purpose to determine work places.



```{r determine_place}
# Use calculated distances (rather than place names) to determine trip end
# place type (H, W, S(chool), O)
calc_dist <- function(x1, y1, x2, y2) {
  df1 <- tibble(lng = x1, lat = y1)
  df2 <- tibble(lng = x2, lat = y2)
  distHaversine(df1, df2)
}

det_home_place <- convert_purpose %>%
  # Home
  left_join(
    hh_combined %>% select(hhid, home_loc_lat, home_loc_lng),
    by = "hhid"
  ) %>%
  mutate(
    orig_home_dist = calc_dist(
      origin_lng, origin_lat, home_loc_lng, home_loc_lat),
    dest_home_dist = calc_dist(
      destination_lng, destination_lat, home_loc_lng, home_loc_lat),
    o_place = ifelse(orig_home_dist < 100, "H", "O"),
    o_place = ifelse(is.na(o_place), "O", o_place),
    o_place = ifelse(origin_name == "HOME", "H", o_place),
    d_place = ifelse(dest_home_dist < 100, "H", "O"),
    d_place = ifelse(is.na(d_place), "O", d_place),
    d_place = ifelse(destination_name == "HOME", "H", d_place)
  ) %>%
  select(-home_loc_lat, -home_loc_lng) %>%
  # correct home activities that don't happen at home
  mutate(
    o_activity = ifelse(o_activity == "H" & o_place != "H", "O", o_activity),
    d_activity = ifelse(d_activity == "H" & d_place != "H", "O", d_activity)
  )

school_loc <- per_combined %>%
  select(hhid, personid, school_address, school_lat, school_lng) %>%
  left_join(
    hh_combined %>% select(hhid, home_loc_address),
    by = "hhid"
  ) %>%
  filter(!is.na(school_lat) & home_loc_address != school_address) %>%
  select(-hhid, -home_loc_address)

det_sch_place <- det_home_place %>%
  left_join(school_loc, by = "personid") %>%
  mutate(
    orig_sch_dist = calc_dist(origin_lng, origin_lat, school_lng, school_lat),
    dest_sch_dist = calc_dist(
      destination_lng, destination_lat, school_lng, school_lat),
    o_place = case_when(
      is.na(orig_sch_dist) ~ o_place,
      o_place == "H" ~ "H",
      # orig_sch_dist < 100 ~ "S",
      origin_address == school_address ~ "S",
      TRUE ~ o_place
    ),
    d_place = case_when(
      is.na(dest_sch_dist) ~ d_place,
      d_place == "H" ~ "H",
      # dest_sch_dist < 100 ~ "S",
      destination_address == school_address ~ "S",
      TRUE ~ d_place
    ),
    # Guess school places for students with missing school loc info
    o_place = ifelse(
      student_type != "" &
        is.na(school_address) &
        o_activity %in% c("PreK", "K12", "U") &
        o_place == "O" &
        str_detect(origin_name, "school"),
      "S", o_place
    ),
    d_place = ifelse(
      student_type != "" &
        is.na(school_address) &
        d_activity %in% c("PreK", "K12", "U") &
        d_place == "O" &
        str_detect(origin_name, "school"),
      "S", d_place
    )
  ) %>%
  select(-school_lng, -school_lat)

# Determining work place is harder. People have multiple jobs, and some jobs
# don't have fixed addresses. Rely on their reported activity.
det_work_place <- det_sch_place %>%
  mutate(
    o_place = case_when(
      o_place == "H" ~ "H",
      o_activity == "W" ~ "W",
      TRUE ~ o_place
    ),
    d_place = case_when(
      d_place == "H" ~ "H",
      d_activity == "W" ~ "W",
      TRUE ~ d_place
    )
  ) %>%
  relocate(o_place, .before = o_activity) %>%
  relocate(d_place, .after = o_place)
```

### Tour formation

While the TRMG2 is a trip-based model, the production rates and other estimated
behavior can still make use of tour information to improve predictive power. At
the same time, the trip-based formulation means that tour formation is much
simpler. Rather than requiring detailed tour pattern information to support
coordinated activity patterns within a household, tours can be classified
simply as work or non-work.

For each person in the survey, a tour begins with their first trip and ends
when they return home. A second tour starts if they leave home again. A tour is
classified as a work tour if the traveler has a work activity during the tour.
If they do not, then the tour is classified as non-work.

There is a third tour type ("home"), that handles cases where the survey reports
trips within the home. For example, when a respondent reports being at home and
then working from home, the survey includes that as a trip. Home tours also
capture things like recreational walks that start and end at the home. For the
purpose of travel modeling, these home tours are ignored.

```{r tour_formation}
determine_tours <- det_work_place %>%
  group_by(personid) %>%
  mutate(
    new_tour = ifelse(row_number() == 1, 1, 0),
    new_tour = ifelse(o_place == "H", 1, new_tour),
    tour_num = cumsum(new_tour)
  ) %>%
  group_by(personid, tour_num) %>%
  mutate(
    tour_anchor = ifelse(d_activity == "W", "W", ""),
    tour_anchor = ifelse(o_place == "H" & d_place == "H", "H", tour_anchor),
    tour_type = case_when(
      "W" %in% tour_anchor ~ "W",
      "H" %in% tour_anchor ~ "H",
      "S" %in% tour_anchor ~ "S",
      TRUE ~ "O"
    ),
    o_party_size_change = party_size - lag(party_size),
    d_party_size_change = lead(party_size) - party_size,
    o_act_dur = departure_time - lag(arrival_time),
    d_act_dur = lead(departure_time) - arrival_time
  )
```

### Problems

```{r}
num_sch_problems <- determine_tours %>%
  mutate(
    sch_flag = case_when(
      o_activity == "SCH?" ~ 1,
      d_activity == "SCH?" ~ 1,
      TRUE ~ 0
    )
  ) %>%
  filter(sch_flag == 1) %>%
  nrow()
```

```{r clean_sch_probs}
# this checks to see if anyone in the trip had school as origin/dest place.
# Drop off trips with school trips often take longer (e.g. waiting for the
# bus or carpool).
s_check <- determine_tours %>%
  ungroup() %>%
  unite(
    key, hhid, origin_address, departure_time, destination_address, arrival_time
  ) %>%
  select(key, o_place, d_place) %>%
  group_by(key) %>%
  mutate(
    o_s_check = ifelse("S" %in% o_place, 1, 0),
    d_s_check = ifelse("S" %in% d_place, 1, 0)
  ) %>%
  summarize(
    o_s_check = max(o_s_check),
    d_s_check = max(d_s_check)
  )

# Use logic to identify pu/do trips. Many of the SCH? activities are actually
# pu/do.
fix_pudo <- determine_tours %>%
  unite(
    key, hhid, origin_address, departure_time, destination_address, arrival_time,
    remove = FALSE
  ) %>%
  left_join(s_check, by = "key") %>%
  mutate(
    # flag trips where the logic identifies DO before changing purpose
    d_do_flag = case_when(
      d_party_size_change >= 0 ~ 0,
      d_act_dur <= 10 ~ 1,
      d_act_dur <= 25 & d_s_check == 1 ~ 1,
      TRUE ~ 0
    ),
    d_activity = ifelse(d_do_flag == 1, "DO", d_activity),
    o_activity = case_when(
      is.na(lag(d_do_flag)) ~ o_activity,
      lag(d_do_flag) == 1 ~ "DO",
      TRUE ~ o_activity
    ),
    d_pu_flag = case_when(
      d_party_size_change <= 0 ~ 0,
      d_act_dur <= 10 ~ 1,
      d_act_dur <= 25 & o_s_check == 1 ~ 1,
      TRUE ~ 0
    ),
    d_activity = ifelse(d_pu_flag == 1, "PU", d_activity),
    o_activity = case_when(
      is.na(lag(d_pu_flag)) ~ o_activity,
      lag(d_pu_flag) == 1 ~ "PU",
      TRUE ~ o_activity
    ),
    # Differentiate PU/DO at school from other
    d_activity = ifelse(
      d_activity %in% c("PU", "DO") & d_s_check == 1,
      paste0(d_activity, "S"), d_activity
    ),
    o_activity = case_when(
      is.na(lag(d_activity)) ~ o_activity,
      lag(d_activity) %in% c("PUS", "DOS") ~ paste0(o_activity, "S"),
      TRUE ~ o_activity
    )
  )

fix_remaining_sch <- fix_pudo %>%
  mutate(
    o_activity = ifelse(o_activity == "SCH?" & o_place == "H", "H", o_activity),
    o_activity = ifelse(o_activity == "SCH?", "O", o_activity),
    d_activity = ifelse(d_activity == "SCH?" & d_place == "H", "H", d_activity),
    d_activity = ifelse(d_activity == "SCH?", "O", d_activity),
  )
```

```{r clean_bus_probs}
fix_bus <- fix_remaining_sch %>%
  mutate(
    bus_flag = ifelse(
      str_detect(tolower(destination_name), "bus") &
        !str_detect(tolower(destination_name), "business") &
        mode != lead(mode) &
        d_activity %in% c("PreK", "K12", "U"),
      1, 0
    ),
    d_activity = ifelse(bus_flag == 1, "X", d_activity),
    o_activity = case_when(
      is.na(lag(d_activity)) ~ o_activity,
      lag(d_activity) == "X" ~ "X",
      TRUE ~ o_activity
    )
  )


```

There are `r num_sch_problems` trip records where a school activity purpose is
recorded for people who are not students.

- `r num_sch_problems` trips recorded with school activity for non-students.
- `r num_missing_joint_trips` trips that reference a household member that not also have that trip listed.

Numerous problems detected:

- School activities reported by non-students (1000 trips)
  - Pickup/Dropoff coded as school (300 origins / 300 origins)
  - Bus stops coded as school rather than mode transfer (70 trips)
- Missing joint trips (1604). E.g., a trip references two hh members but second
  member doesn't have that trip
- Kids marked as homeschool but go to traditional school on travel day for
  school activity (30)
- Students without school address in person file (195)
- Home activities reported at non-home locations (e.g. "In-laws house") (2,147)
  - This is after we extended the home definition to be 100m around the home (for things like 'front yard' and 'tractor shed')

## Income imputation

The survey provides income at two levels of aggregation: 10 categories or 5. RSG
imputed incomes for the broad category in order to use ACS income information
during survey expansion. The imputation:

  1. Further stratified category 5 (\$100k+) into those above and below \$150k.
  2. Estimated incomes for those households that did not report.
  
In order to preserve all the records in the survey, the Caliper team used the
imputed, broad categories of income. The table below shows the count of samples
by the original and imputed categories.

```{r}
temp_broad <- bind_rows(
  remove_na_2016 %>%
    select(hh_income_broad),
  remove_na_2018 %>%
    select(hh_income_broad)
) %>%
  group_by(hh_income_broad) %>%
  summarize(count = n()) %>%
  ungroup() %>%
  rename(Original = hh_income_broad, count1 = count)

df <- bind_cols(
  temp_broad,
  hh_combined %>%
    group_by(hh_income_broad) %>%
    summarize(count = n()) %>%
    ungroup() %>%
    rename(Imputed = hh_income_broad, count2 = count)
)
colnames(df) <- c("Original", "Count", "Imputed", "Count")
df$Original <- c(
  "Under $25,000",
  "$25,000-$49,999",
  "$50,000-$74,999",
  "$75,000-$99,999",
  "$100,000 or more",
  "Prefer not to answer"
)
df$Imputed <- c(
  "Under $25,000",
  "$25,000-$49,999",
  "$50,000-$74,999",
  "$75,000-$99,999",
  "$100,000-$149,999",
  "$150,000 or more"
)

df %>%
  kable() %>%
  kable_styling()
```

A full discussion of their income imputation can be found in the survey
documentation. They used ordinal logistic regression to predict income
categories based on household features like size and number of households (among
others).



```{r write tables, eval=FALSE}
# This treatment prevents the output files from being changed unintentionally
# while knitting, testing, etc.
if {1 = 0} {
  write_csv(
    hh_combined,
    "data/output/_PRIVATE/survey_processing/hh_processed.csv"
  )
  write_csv(
    per_combined,
    "data/output/_PRIVATE/survey_processing/per_processed.csv"
  )
  write_csv(
    determine_tours,
    "data/output/_PRIVATE/survey_processing/trips_processed.csv"
  )
}
```

