---
title: "TAZs and SE Data"
author: "Caliper Corporation"
date: "12/8/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(dplyr.summarise.inform = FALSE)
options(scipen = 999)

library(tidyverse)
library(sf)
library(leaflet)
```

## TAZ edits

The zonal system for TRMG2 was inherited from the existing model, and Caliper
performed a brief review and recommended additional zone splits. These
recommendations were focused on addressing two concerns:

1. Breaking up zones with large employment and households (in base or future).
2. Adding granularity for universities.

Large zones can create traffic loading problems by generating lots of trips and
loading them on the network in a small number of locations. Breaking these into
smaller zones allows for more-even loading of traffic and avoids spikes in
congestion.

A major concern for universities at the outset is getting transit loading
correct. Larger zones capture more trips as "intra-zonal" meaning they never
leave the zone and load onto the network. For universities, where trips between
classes have a significant impact on transit ridership, it is important that
these trips travel between zones. In addition, the added granularity gives more
accurate measurements of non-motorized travel times, which is important for the
university market.

Caliper made several presentations of proposed changes and took feedback from
the stakeholders concerning zone edits. The map below shows the final edits.
Scroll/Zoom the map below to review them, and use the layers control box
to turn the old and new zone layers off.

```{r}
taz_old <- sf::read_sf("data/input/tazs/tazs 2020-11-02.shp") %>%
  st_transform(crs = st_crs('+proj=longlat +datum=WGS84'))
taz_new <- sf::read_sf("data/input/tazs/tazs 2020-12-08.shp") %>%
  st_transform(crs = st_crs('+proj=longlat +datum=WGS84'))
```

```{r}
leaflet() %>%
  addProviderTiles("Stamen.TonerLite") %>%
  addPolygons(
    data = taz_new,
    weight = 1,
    fill = FALSE,
    color = "red",
    group = "New"
  ) %>%
  addPolygons(
    data = taz_old,
    weight = 1,
    fill = FALSE,
    group = "Old"
  ) %>%
  addLayersControl(overlayGroups = c("New", "Old")) %>%
  setView(-78.672682, 35.784927, zoom = 12)
```

### Network updates

For each new TAZ, a corresponding centroid node and appropriate centroid
connectors were added to the roadway network.
  
### Allocating existing SE data

*Placeholder*

For now, the SE data is allocated into the new zones using the percent area.

```{r}
se_data <- read_csv(
  "data/input/se_data/se_2016.csv",
  col_types = cols(
    .default = col_double(),
    County = col_character(),
    TRM_6.2_Fill = col_logical()
  )
)

calc_percents <- se_data %>%
  left_join(taz_new %>% st_drop_geometry(), by = c("TAZ_NG" = "ORIG_ID")) %>%
  group_by(TAZ_NG) %>%
  mutate(pct = AREA / sum(AREA)) %>%
  mutate(across(
    c(HH, HH_POP, Stud_GQ:Retail),
    ~round(.x * pct, 0)
  )) %>%
  mutate(
    TAZ_NG = ifelse(is.na(ID), TAZ_NG, ID),
    County = ifelse(County == "External", "External", "Internal")
  ) %>%
  select(
    TAZ = TAZ_NG, Type = County, HH:Retail, K12 = ENROLLMENT,
    ParkCostW = PrkCosW1, ParkCostO = PrkCosO1, ParkCostU = PrkCosU1, 
    ADT:PCTCVEE
  ) %>%
  arrange(TAZ)

# Throw error if total households are off by more than rounding error
hh_ratio = sum(calc_percents$HH, na.rm = TRUE) / 
  sum(se_data$HH, na.rm = TRUE)
if (abs(hh_ratio - 1) > .01) stop(
  "Disaggregation caused significant change in total households"
)
```

## Creating new SE data

The TRMG2 model will make use of some new zonal attributes:

- Median Income
- Workers
- Vehicles

Caliper used the 2014-2018 5-year ACS data to populate these new attributes.

### Income

```{r load census data, include=FALSE}
source("R/load_census_data.R")
shapes <- load_census_data()

acs_bg <- shapes$acs_bg
acs_tract <- shapes$acs_tract
dec_bg <- shapes$dec_bg

taz <- st_transform(taz_new, st_crs(acs_bg)) %>% st_make_valid()
```

```{r, include=FALSE}
# regional median income
# source: https://data.census.gov/cedsci/table?q=median%20income&g=0500000US37063,37135,37183&tid=ACSST1Y2019.S1901&hidePreview=true
reg_med_inc <- 65317

inc_centroids <- taz %>%
  st_point_on_surface() %>%
  st_join(acs_bg %>% select(inc_med)) %>%
  as.data.frame() %>%
  select(ID, inc_med)
  
num_missing <- nrow(inc_centroids %>% filter(is.na(inc_med)))
add_income_taz <- taz %>%
  left_join(inc_centroids, by = "ID") %>%
  mutate(inc_med = ifelse(is.na(inc_med), reg_med_inc, inc_med)) %>%
  rename(MEDIANINC = inc_med)

add_income_se <- calc_percents %>%
  left_join(
    add_income_taz %>%
      select(ID, MEDIANINC) %>%
      st_drop_geometry(),
    by = c("TAZ" = "ID")
  ) %>%
  relocate(MEDIANINC, .before = MEANINC) %>%
  select(-MEANINC)
```

Due to the skewed nature of income distributions, median income is a better
approximation of the average than the mean. This information was appended to the
socio-economic (SE) data from the Census, but income information is only
available at the block group geography level. As a result, all TAZs within a
block group have the same median income. In addition, income measures were
suppressed by the Census for `r num_missing` block groups. For TAZs in these
block groups, the regional median income of `r paste0('$',formatC(reg_med_inc,
big.mark=',', format = 'd'))` was used.

### Workers

```{r, include=FALSE}
calc_tot_workers <- acs_tract %>%
  mutate(
    # 3.1 acknowledges that 3+worker households have an average higher than 3,
    # but the exact average is unknown.
    wrk_tot = wrk1 + 2 * wrk2 + 3.1 * wrk3,
    avg_wrk = wrk_tot / (hh_tot + .0000001)
  )

wrk_centroids <- taz %>%
  st_point_on_surface() %>%
  st_join(calc_tot_workers %>% select(avg_wrk)) %>%
  as.data.frame() %>%
  select(ID, avg_wrk)
  
add_workers_taz <- add_income_taz %>%
  left_join(wrk_centroids, by = "ID")

add_workers_se <- add_income_se %>%
  left_join(
    add_workers_taz %>%
      select(ID, avg_wrk) %>%
      st_drop_geometry(),
    by = c("TAZ" = "ID")    
  ) %>%
  mutate(Workers = round(avg_wrk * HH, 0)) %>%
  select(-avg_wrk) %>%
  relocate(Workers, .after = MEDIANINC)
```

Estimates of workers are only available from the ACS at the tract level. It is
also important to note that ACS tables that count total workers cannot be used,
because they include workers living in group quarters. Instead, total workers in
households must be imputed from tables that count households by number of
workers.

Without more-disaggregate data to inform the process, the following steps were
taken to assign workers to TAZs:

  1. Calculate the average workers per household for the ACS tract.
  2. Use that average for each TAZ within the tract.
  
In this way, workers were allocated within each tract proportionally based
on the number of households.

### Vehicles

```{r, include=FALSE}
reg_avg_veh <- sum(acs_bg$veh_tot, na.rm = TRUE) / sum(acs_bg$hh_tot, na.rm = TRUE)

acs_bg <- acs_bg %>%
  mutate(avg_veh = veh_tot / hh_tot)

veh_centroids <- taz %>%
  st_point_on_surface() %>%
  st_join(acs_bg %>% select(avg_veh)) %>%
  as.data.frame() %>%
  select(ID, avg_veh)

add_veh_taz <- add_workers_taz %>%
  left_join(veh_centroids, by = "ID") %>%
  mutate(avg_veh = ifelse(is.na(avg_veh), reg_avg_veh, avg_veh))
  
add_veh_se <- add_workers_se %>%
  left_join(
    add_veh_taz %>%
      select(ID, avg_veh) %>%
      st_drop_geometry(),
    by = c("TAZ" = "ID")    
  ) %>%
  mutate(
    Vehicles = round(HH * avg_veh, 0)
  ) %>%
  select(-avg_veh) %>%
  relocate(Vehicles, .after = Workers)
```

Vehicle data is available from the ACS at the block group level. Unlike workers,
total vehicles can be retrieved for just households; however, the lack of
information at the block level means that all TAZs within the same block group
are assigned the same number of vehicles per household. Also, some block groups
had suppressed vehicle information and the regional average of 
`r round(reg_avg_veh, 2)` was used.

```{r}
if (1 == 0) {
  write_csv(
    add_veh_se, "data/output/se_data/se_2016.csv",
    na = "0"
  )
}
```