---
title: "Time of Day, Directionality, and Occupancy"
author: "Caliper Corporation"
date: "February 8, 2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warnings = FALSE)
options(dplyr.summarise.inform = FALSE)
options(scipen = 999)

library(tidyverse)
library(data.table)
library(chron)
library(plotly)
library(knitr)
library(kableExtra)
library(scales)
```

# Introduction
This page details the steps taken to determine the most appropriate time period
definitions for the Triangle Regional Model (TRMG2). Caliper used the processed
household survey to perform the analysis.

```{r read survey, include = FALSE}
# This file is created by the Generation-Resident.Rmd
tripTbl <- read_csv("data/output/_PRIVATE/survey_processing/trips_processed.csv")
```

# Peak period determination

*(To skip to the final period definitions, click [here](#final_defs).)*

Peak period determination is based on an analysis of the trips in motion
throughout the day. First, the day is broken up into 15-minute increments. A
trip from the household survey is said to be "in motion" if any portion of the
trip occurs within the 15-minute bin. As a result, a single trip can be counted
in multiple bins. Determining the peak period considers the distribution of all
trips as well as the distribution of trips on work tours individually. The chart
below shows these two distributions, which look as expected with the AM peak
being shorter and more condensed than the PM. 

```{r}
bin_size = 15 # minutes

# Vector of time by 15 minutes
v_time <- seq(0, 24, by = bin_size / 60) * 60

# Determine how many trips (all purposes) fall in each bin
v_all <- c()
for (i in 1:length(v_time)) {
  v_all[i] <- sum(
    tripTbl$trip_weight_combined[
      tripTbl$departure_time < v_time[i + 1] & 
        tripTbl$arrival_time >= v_time[i]
    ],
    na.rm = TRUE
  )
}

# Determine how many trips (work purpose) fall in each bin
work_trips <- tripTbl %>%
  mutate(
    work_trip = ifelse(
      tour_type == "W", #& homebased == "HB" & purp_scheme == "W",
      # (o_activity == "W" | d_activity == "W") &
      #   (o_activity == "H" | d_activity == "H"),
      1, 0
    )
  )

v_wrk <- c()
for (i in 1:length(v_time)) {
  v_wrk[i] <- sum(
    work_trips$trip_weight_combined[
      work_trips$departure_time < v_time[i + 1] & 
        work_trips$arrival_time >= v_time[i] & 
        work_trips$work_trip == 1
    ]
  )
}

timTbl <- tibble(
  time = v_time,
  AllTrips = v_all,
  WrkTrips = v_wrk
)
```

```{r time_conv_formula}
# Formula for converting minutes to chart format (e.g. 1000 for 10:00 am)
chart_time <- function(minutes){
  ((minutes %/% 60) * 100 + (minutes %% 60)) / 100
}
```

```{r, warning=FALSE, message=FALSE}
tbl <- timTbl %>%
  mutate(time = chart_time(time))

plot_ly(
  tbl, x = ~time, y = ~AllTrips, type = "scatter", mode = "lines", name = "All",
  hoverinfo = "text+x", text = ~format(round(AllTrips, 0), big.mark = ",")
) %>%
  add_trace(
    x = ~time, y = ~WrkTrips, type = "scatter", mode = "lines", name = "Work",
    hoverinfo = "text+x", text = ~format(round(WrkTrips, 0), big.mark = ",")
  ) %>%
  layout(
    title = "Distribution of Trips in Motion by Purpose",
    xaxis = list(title = "Time of Day"),
    yaxis = list(title = "Trips in Motion")
  )
```

The next step is to determine the peak hour for all trips and those on work
tours. In the tables below, the AM and PM peak hours by type are presented. At
a minimum, the AM and PM peak periods should contain the respective peak hour
for work trips and all trips. The chart below shows that the peak hour for work
and all trips are exactly the same in the Triangle region based on the surveys 
as is often, but not always, the case.

```{r}
# Create rolling sums to determine peak hour
an = function(n, len) c(seq.int(n), rep(n, len-n))
n = an(60 / bin_size, nrow(timTbl))

roll_sums <- timTbl %>%
  mutate(
    all_trips_in_hour = frollsum(AllTrips, n, na.rm = TRUE, adaptive = TRUE),
    wrk_trips_in_hour = frollsum(WrkTrips, n, na.rm = TRUE, adaptive = TRUE)
  )
```

```{r am_peak_hour}
# AM Peak Hour
am_tbl <- roll_sums %>%
  filter(time < 720) # before noon
# for work trips
am_wrk_peak <- am_tbl$time[
  which(am_tbl$wrk_trips_in_hour == max(am_tbl$wrk_trips_in_hour))
  ]
# for all trips
am_all_peak <- am_tbl$time[
  which(am_tbl$all_trips_in_hour == max(am_tbl$all_trips_in_hour))
  ]

# Convert minutes to times for table and prose
am_wrk_peak_from <- chron(times. = c((am_wrk_peak - 60) / 1440))
am_wrk_peak_to <- chron(times. = c(am_wrk_peak / 1440))
am_all_peak_from <- chron(times. = c((am_all_peak - 60) / 1440))
am_all_peak_to <- chron(times. = c(am_all_peak / 1440))
am_min_from <- chron(times. = c((min(am_wrk_peak, am_all_peak) - 60) / 1440))
am_max_to <- chron(times. = c(max(am_wrk_peak, am_all_peak) / 1440))
  
tibble(
  `Trip Purpose` = c("Work", "All"),
  `Peak Start` = c(am_wrk_peak_from, am_all_peak_from),
  `Peak End` = c(am_wrk_peak_to, am_all_peak_to)
) %>%
  # kable2(caption = "AM Peak Hour by Purpose")
  kable() %>%
  kable_styling(full_width = FALSE)

```

```{r pm_peak_hour}
# PM Peak Hour
pm_tbl <- roll_sums %>%
  filter(time > 720) # after noon
# for wrk trips
pm_wrk_peak <- pm_tbl$time[
  which(pm_tbl$wrk_trips_in_hour == max(pm_tbl$wrk_trips_in_hour))
  ]
# for all trips
pm_all_peak <- pm_tbl$time[
  which(pm_tbl$all_trips_in_hour == max(pm_tbl$all_trips_in_hour))
  ]

# Convert minutes to times for table and prose
pm_wrk_peak_from <- chron(times. = c((pm_wrk_peak - 60) / 1440))
pm_wrk_peak_to <- chron(times. = c(pm_wrk_peak / 1440))
pm_all_peak_from <- chron(times. = c((pm_all_peak - 60) / 1440))
pm_all_peak_to <- chron(times. = c(pm_all_peak / 1440))
pm_min_from <- chron(times. = c((min(pm_wrk_peak, pm_all_peak) - 60) / 1440))
pm_max_to <- chron(times. = c(max(pm_wrk_peak, pm_all_peak) / 1440))
  
tibble(
  `Trip Purpose` = c("Work", "All"),
  `Peak Start` = c(pm_wrk_peak_from, pm_all_peak_from),
  `Peak End` = c(pm_wrk_peak_to, pm_all_peak_to)
) %>%
  # kable2(caption = "PM Peak Hour by Purpose")
  kable() %>%
  kable_styling(full_width = FALSE)

```

```{r peak_hour_chart}
plotly_tbl <- timTbl %>%
  mutate(time = chart_time(time))

plot_ly(
  plotly_tbl,
  x = ~time, y = ~AllTrips, type = "scatter", mode = "lines", name = "All",
  hoverinfo = "text+x", text = ~format(round(AllTrips, 0), big.mark = ",")
) %>%
  add_trace(
    x = ~time, y = ~WrkTrips, type = "scatter", mode = "lines", name = "Work",
    hoverinfo = "text+x", text = ~format(round(WrkTrips, 0), big.mark = ",")
  ) %>%
  add_trace(
    name = "Peak Hours",
    hoverinfo = "x",
    # Vector of x positions (may need individual adjustment)
    x = c(
      chart_time(am_wrk_peak - 10),
      chart_time(pm_wrk_peak - 35)
      ),
    # Vector of y positions
    y = rep(max(plotly_tbl$AllTrips) * 1.02, 2),
    text = c("Work and All", "Work and All"),
    mode = "text"
  ) %>%
  layout(
    title = "Peak Hours",
    xaxis = list(title = "Time of Day"),
    yaxis = list(title = "Trips in Motion"),
    shapes = list(
      # Work AM Peak Hour
      list(type = "rect",
           fillcolor = "blue", line = list(color = "blue"), opacity = .3,
           x0 = chart_time(am_wrk_peak - 60),
           x1 = chart_time(am_wrk_peak),
           y0 = 0, y1 = max(plotly_tbl$AllTrips)
      ),
      # All AM Peak Hour
      list(type = "rect",
           fillcolor = "green", line = list(color = "green"), opacity = .3,
           x0 = chart_time(am_all_peak - 60),
           x1 = chart_time(am_all_peak),
           y0 = 0, y1 = max(plotly_tbl$AllTrips)
      ),
      # Work PM Peak Hour
      list(type = "rect",
           fillcolor = "blue", line = list(color = "blue"), opacity = .3,
           x0 = chart_time(pm_wrk_peak - 60),
           x1 = chart_time(pm_wrk_peak),
           y0 = 0, y1 = max(plotly_tbl$AllTrips)
      ),
      # All PM Peak Hour
      list(type = "rect",
           fillcolor = "green", line = list(color = "green"), opacity = .3,
           x0 = chart_time(pm_all_peak - 60),
           x1 = chart_time(pm_all_peak),
           y0 = 0, y1 = max(plotly_tbl$AllTrips)
      )
    )
  )
```


```{r period_definition}
#If desired, manually override the final period definitions in the variables
#below. Use chron(times. = c("15:30:00")) to specify 3:30pm

am_peak_from <- am_min_from
# am_peak_to <- am_max_to
am_peak_to <- chron(times. = c("9:00:00"))

# pm_peak_from <- pm_min_from
# pm_peak_to <- pm_max_to
pm_peak_from <- chron(times. = c("15:30:00"))
pm_peak_to <- chron(times. = c("18:15:00"))
```

## Final Period Definitions {#final_defs}

Based on the trips in motion profile, there are four distinct periods of the
day. AM and PM peaks have the highest intensity of trip making, followed by
mid-day, with the overnight period containing the fewest trips in motion. 
Boundaries for these periods are defined such that the variance of trips in
motion within periods is smallest while variance between periods is largest.

The final period definitions are shown in the table below including the mid-day
(MD) period. The remaining hours of the day are captured in the night (NT)
period. In the chart, the final period is represented by the gray rectangle.
```{r final_per_defs}
period_tbl <- tibble(
  Period = c("AM", "MD", "PM", "NT"),
  From = c(am_peak_from, am_peak_to, pm_peak_from, pm_peak_to),
  To = c(am_peak_to, pm_peak_from, pm_peak_to, am_peak_from)
) %>%
  mutate(
    Hours = hours(To - From) + minutes(To - From) / 60
  )

period_tbl %>%
  # kable2(caption = "Final Period Definitions")
  kable() %>%
  kable_styling(full_width = FALSE)
```

```{r peak_period_chart}
# Function used to calculate the x location of the period label
text_loc <- function (chron1, chron2) {
  start <- as.numeric(min(chron1, chron2))
  stop <- as.numeric(max(chron1, chron2))
  midpoint <- ((stop - start) / 2 + start) * 1440
  midpoint <- chart_time(midpoint)
  return(midpoint)
}

plot_ly(
  plotly_tbl,
  x = ~time, y = ~AllTrips, type = "scatter", mode = "lines", name = "All",
  hoverinfo = "text+x", text = ~format(round(AllTrips, 0), big.mark = ",")
) %>%
  add_trace(
    x = ~time, y = ~WrkTrips, type = "scatter", mode = "lines", name = "Work",
    hoverinfo = "text+x", text = ~format(round(WrkTrips, 0), big.mark = ",")
  ) %>%
  add_trace(
    name = "Period",
    hoverinfo = "x",
    # Vector of x positions
    x = c(
      text_loc(0, am_peak_from),                        # first NT
      text_loc(am_peak_from, am_peak_to),               # AM
      text_loc(am_peak_to, pm_peak_from),               # MD
      text_loc(pm_peak_from, pm_peak_to),               # PM
      text_loc(pm_peak_to, chron(times. = "23:59:59"))  # second NT
      ),
    # Vector of y positions
    y = rep(max(plotly_tbl$AllTrips) * 1.06, 5),
    text = c("NT", "AM", "MD", "PM", "NT"),
    mode = "text"
  ) %>%
  layout(
    title = "Peak Periods",
    xaxis = list(title = "Time of Day"),
    yaxis = list(title = "Trips in Motion"),
    shapes = list(
      # AM Peak Period
      list(type = "rect",
           fillcolor = "gray", line = list(color = "black"), opacity = .3,
           x0 = chart_time(as.numeric(am_peak_from) * 1440),
           x1 = chart_time(as.numeric(am_peak_to) * 1440),
           y0 = 0, y1 = max(plotly_tbl$AllTrips) * 1.04
      ),
      # AM Peak Period
      list(type = "rect",
           fillcolor = "gray", line = list(color = "black"), opacity = .3,
           x0 = chart_time(as.numeric(pm_peak_from) * 1440),
           x1 = chart_time(as.numeric(pm_peak_to) * 1440),
           y0 = 0, y1 = max(plotly_tbl$AllTrips) * 1.04
      ),
      # Work AM Peak Hour
      list(type = "rect",
           fillcolor = "blue", line = list(color = "blue"), opacity = .3,
           x0 = chart_time(am_wrk_peak - 60),
           x1 = chart_time(am_wrk_peak),
           y0 = 0, y1 = max(plotly_tbl$AllTrips)
      ),
      # All AM Peak Hour
      list(type = "rect",
           fillcolor = "green", line = list(color = "green"), opacity = .3,
           x0 = chart_time(am_all_peak - 60),
           x1 = chart_time(am_all_peak),
           y0 = 0, y1 = max(plotly_tbl$AllTrips)
      ),
      # Work PM Peak Hour
      list(type = "rect",
           fillcolor = "blue", line = list(color = "blue"), opacity = .3,
           x0 = chart_time(pm_wrk_peak - 60),
           x1 = chart_time(pm_wrk_peak),
           y0 = 0, y1 = max(plotly_tbl$AllTrips)
      ),
      # All PM Peak Hour
      list(type = "rect",
           fillcolor = "green", line = list(color = "green"), opacity = .3,
           x0 = chart_time(pm_all_peak - 60),
           x1 = chart_time(pm_all_peak),
           y0 = 0, y1 = max(plotly_tbl$AllTrips)
      )
    )
  )
```


The table below shows the average 15-minute trips in motion during each period.
These period definitions accurately capture the disparate characteristics of
each.

```{r}
am_from_min <- as.numeric(am_peak_from) * 1440
am_to_min <- as.numeric(am_peak_to) * 1440
pm_from_min <- as.numeric(pm_peak_from) * 1440
pm_to_min <- as.numeric(pm_peak_to) * 1440

avg_tim <- timTbl %>%
  mutate(tod = case_when(
    time <= am_from_min ~ "NT",
    time <= am_to_min ~ "AM",
    time <= pm_from_min ~ "MD",
    time <= pm_to_min ~ "PM",
    TRUE ~ "NT"
  )) %>%
  group_by(tod) %>%
  summarize(mean = mean(AllTrips)) %>%
  mutate(tod = factor(
    tod, levels = c("AM", "MD", "PM", "NT"), ordered = TRUE
  )) %>%
  arrange(tod)

avg_tim %>%
  mutate(mean = round(mean, 0)) %>%
  rename(`Time of Day` = tod, `Mean Trips in Motion` = mean) %>%
  kable() %>%
  kable_styling(full_width = FALSE)
```

# Time of day factors
With the periods defined, the percent of trips produced by purpose and time of
day must be determined. These factors are calculated based on the departure
time of each trip rather than using the trips in motion table. This ensures
that each trip is only counted once, regardless of duration, and that it is
included in the right period. These factors are used to divide the daily person
trips created during trip generation into time of day (by purpose).

Directionality factors are calculated separately in the next section. As a
reference, the time period definition is shown again first followed by the TOD
factors.

```{r}
period_tbl %>%
  kable() %>%
  kable_styling(full_width = FALSE)
```

```{r calc_tod_trip_factors}
tod_factors <- tripTbl %>%
  filter(tour_type != "H", homebased == "HB") %>%
  group_by(tod, trip_type) %>%
  summarise(Trips = sum(trip_weight_combined)) %>%
  group_by(trip_type) %>%
  mutate(
    factor = round(Trips / sum(Trips), 3),
    # ensure unity
    temp = ifelse(tod == "NT", 0, factor),
    factor = ifelse(tod == "NT", 1 - sum(temp), factor)
  ) %>%
  select(trip_type, tod, factor) %>%
  arrange(trip_type, tod)

tod_factors %>%
  ungroup() %>%
  mutate(
    factor = percent(factor, accuracy = .1),
    tod = factor(tod, levels = c("AM", "MD", "PM", "NT"), ordered = TRUE)
  ) %>%
  # filter(trip_type %in% c("W_HB_W_All", "N_HB_K12_All")) %>%
  spread(key = tod, value = factor) %>%
  rename(`Trip Type` = trip_type) %>%
  kable() %>%
  kable_styling(full_width = FALSE)

# Create a description field for the table
desc <- vector(mode = "character", length = nrow(tod_factors))
desc[1] = "Used to split daily resident productions"
desc[2] = "into period-specific values"
tod_factors$Description <- desc
```

# Directionality factors
For most steps in a trip-based model, the trips are said to be in
"Production/Attraction" (PA) format. This simply means that all home-based
trips start at home and end somewhere else. This simplification of reality is
done for a number of reasons in the model, but is not actually how travel
occurs. As a result, before highway trip assignment can take place, the PA
format must be converted to "Origin/Destination" (OD) format. In this format, a
trip from work to home starts at work and ends at home.

This conversion is accomplished using factors stratified by time of day and trip
type. As an example, the majority of W_HB_W_All (i.e. HBW) trips in the AM start
at home and end at work. As a result, the PA-to-OD factor would be above 0.5. In
the PM period, this trend is usually reversed, and the PA factor would be less
than 0.5.

These factors can be calculated directly from the survey by comparing the number
of home-based trips that start at home to the number home-based trips that start
elsewhere. By definition, non-home-based trips are the same in either format;
their PA factors are set to .5. The same treatment is applied to commercial
vehicles, trucks, and external trips.

Directionality factors are applied after the distribution and modal
models. At that point, trip types are collapsed due to sample size
limitations in the household survey. As a result, the factors must be estimated
using the matching, more-aggregate purposes.

*TODO: Calculate these after aggregate models are specified*

```{r calc_pa_factors}
# # Calculate a pa field
# tripTbl <- tripTbl %>%
#   mutate(pa = ifelse(dest_blk == home_blk & orig_blk != home_blk, "AP", "PA"))
# 
# pa_tbl <- tripTbl %>%
#   group_by(Purpose = final_purp_d, Period = tod, pa) %>%
#   summarise(Trips = sum(hhrkwt0)) %>%
#   group_by(Purpose, Period) %>%
#   mutate(
#     Value = round(Trips / sum(Trips), 3),
#     # Set NHB purposes to .5
#     Value = ifelse(
#       # Check for first letter = "N" or "n"
#       grepl("^N", Purpose, ignore.case = TRUE),
#       .5,
#       Value
#     )
#   ) %>%
#   filter(pa == "PA") %>%
#   select(Purpose, Period, Value) %>%
#   ungroup() %>%
#   mutate(Purpose = factor(
#     Purpose, levels = c("HBW", "HBO", "HBSC", "HBSH", "NHB"), ordered = TRUE
#   )) %>%
#   arrange(Purpose)
# 
# pa_tbl %>%
#   spread(key = Period, value = Value) %>%
#   kable2(digits = 2, caption = "PA Factors by Purpose and Time of Day")
```

```{r add other purps and desc, warning=FALSE}
# # Add rows for NHBNR, CV, SUT, MUT, and IEEI. They are all 50/50.
# v_types <- c("NHBNR", "CV", "SUT", "MUT", "IEEI")
# for (type in v_types){
#   template <- pa_tbl %>%
#     filter(Purpose == "NHB") %>%
#     mutate(Purpose = type)
#   
#   pa_tbl <- bind_rows(pa_tbl, template)
# }
# pa_tbl <- pa_tbl %>%
#   mutate(Purpose = factor(
#     Purpose, levels = c(
#       "HBW",
#       "HBO",
#       "HBSC",
#       "HBSH",
#       "NHB",
#       v_types
#     ), ordered = TRUE
#   )) %>%
#   arrange(Purpose)
# 
# # add description column
# v_desc <- vector(mode = "character", nrow(pa_tbl))
# v_desc[1] <- "Percent of trips in PA direction by purpose and tod"
# v_desc[2] <- "CREATION: this file created by rmarkdown"
# pa_tbl$Description <- v_desc
```

# Vehicle occupancy factors
In addition to applying directionality factors before assignment, the person 
trip matrices are converted to vehicles using occupancy factors. These factors
are calculated from the household survey, which includes information on the
party size of each trip. The rates below meet expectations. The mode categories
used are defined as follows:

- SOV: Single occupancy
- HOV2: 2-person occupancy
- HOV3: 3+ person occupancy

Vehicle occupancy for SOV and HOV2 modes are 1.0 and 2.0, respectively. The
table below shows the occupancy calculated for sr3.

```{r calculate veh occ factors}
occ_tbl <- tripTbl %>%
  filter(tour_type != "H", !is.na(mode_final)) %>%
  filter(mode_final %in% c("sov", "hov2", "hov3")) %>%
  mutate(
    party_size = num_hh_members + num_nonhh,
    mode = mode_final
  ) %>%
  group_by(trip_type, tod, mode) %>%
  mutate(
    Value = round(weighted.mean(party_size, w = trip_weight_combined), 3),
    Samples = n(),
  ) %>%
  # Low samples for some rows leads to high factors (e.g. 4.9). If the sample size
  # is low, use the daily average.
  group_by(trip_type, mode) %>%
  mutate(
    daily_value = round(weighted.mean(party_size, w = trip_weight_combined), 3),
    Value = ifelse(Samples < 30, daily_value, Value)
  ) %>%
  group_by(Purpose = trip_type, Period = tod, Mode = mode) %>%
  slice(1) %>%
  select(Purpose, Period, Mode, Value, Samples) %>%
  ungroup() %>%
  arrange(Purpose, Period, Mode) %>%
  select(Purpose, Mode, Period, Value)

occ_tbl %>%
  spread(key = Period, value = Value) %>%
  filter(Mode == "hov3") %>%
  kable(digits = 2) %>%
  kable_styling(full_width = FALSE)

# TODO: create parameter table

# # Repeat occupancy factors by market segment to match model
# # matrix core names
# occ1 <- occ_tbl %>%
#   filter(Purpose %in% c("HBW", "HBO")) %>%
#   merge(c("v0", "ilvi", "ilvs", "ihvi", "ihvs")) %>%
#   select(Purpose, Segment = y, Mode, Period, Value)
# occ2 <- occ_tbl %>%
#   filter(!(Purpose %in% c("HBW", "HBO"))) %>%
#   mutate(Segment = "all") %>%
#   select(Purpose, Segment, Mode, Period, Value)
# occ_tbl <- bind_rows(occ1, occ2) %>%
#   arrange(Purpose, Segment, Mode, Period)
# 
# 
# # No need to add CV, IX, etc. By not being in the table, it's the same
# # as if they were and had a factor of 1.
# 
# # add description column
# v_desc <- vector(mode = "character", nrow(occ_tbl))
# v_desc[1] <- "Vehicle occupancy rates (per/veh)"
# v_desc[2] <- "CREATION: this file created by TimeOfDay.Rmd and University.Rmd"
# occ_tbl$Description <- v_desc
```


# Period capacity factors
In aggregate period assignment regimes, hourly capacities must be
converted to period capacities in a manner that accurately captures the balance
of congestion. In the real world, the hourly capacity of a road is generally
the same throughout the day; it is the demand that changes. This change in
demand causes congestion levels to change over time. Consider the
stylized example below of a real world street segment.

## Qualitative example

Note that capacity remains fixed while demand changes.

```{r}
real_world <- tibble(
  Time = c("6-7", "7-8", "8-9"),
  Capacity = c(1050, 1050, 1050),
  Demand = c(500, 900, 500)
) %>%
  mutate(Percent = round(Demand / Capacity * 100, 2))

real_world %>% 
  kable() %>%
  kable_styling(full_width = FALSE)

plot_ly(
  real_world,
  x = ~Time, y = ~Capacity, type = "bar", name = "Capacity"
) %>%
  add_trace(x = ~Time, y = ~Demand, type = "bar", name = "Demand") %>%
  layout(
    xaxis = list(title = "Hour of Day"), yaxis = list(title = "Veh per Hour")
  )
```

The easiest way to represent this in a period assignment would be to add up the
total capacity and total demand from each hour.

```{r}
model_period <- tibble(
  Time = c("6-9"),
  Capacity = c(sum(real_world$Capacity)),
  Demand = c(sum(real_world$Demand))
)

model_period %>%
  kable() %>%
  kable_styling(full_width = FALSE)

plot_ly(
  model_period,
  x = ~Time, y = ~Capacity, type = "bar", name = "Capacity"
) %>%
  add_trace(x = ~Time, y = ~Demand, type = "bar", name = "Demand") %>%
  layout(
    xaxis = list(title = "Hour of Day"), yaxis = list(title = "Veh per Hour")
  )
```

In this simple representation, no one experiences congestion. This does not
match the real world graph, which shows that nearly half the vehicles in the
period experience congested conditions. This occurs between 7:00 am and 8:00 am.
This discrepancy between the real world and a naive period representation can
cause numerous errors in the model, including:

- Longer trip lengths
- Incorrect route choices
- Lower diversion to transit
- Incorrect link assignment

The period capacity factor (PCF) corrects for this, and is the inverse of a
similar concept from traffic engineering: the peak hour factor (PHF). In traffic
engineering, the PHF adjusts hourly volume based on the highest 15-minute volume
within the hour. While the PHF is used to adjust demand given a fixed capacity,
the PCF adjusts period capacity based on the highest-volume hour of the period.

Consider our example of a three-hour period from 6:00 am to 9:00 am. If the
actual demand was evenly distributed across the period, the PHF would be .33. 
The PCF, the inverse of the PHF, would be 3. In this scenario, the hourly
capacity would be multiplied by 3, which would lead to the capacity shown in the
table above. Instead, using the actual demand from our example, the PHF would
be:

$$PHF = 900 /(500 + 900 + 500) = 0.474$$

Taking the inverse gives a PCF of:

$$PCF = 1 / .474 = 2.11$$

The period capacity that most-accurately reflects congestion would be:

$$Period Capacity = 1050 * 2.11 = 2215$$

That capacity is shown relative to the period demand in the chart below. This
relationship is a better representation of real world conditions.

```{r}
final_period <- tibble(
  PCF = c("PCF = 3.00", "PCF = 2.11"),
  Demand = rep(sum(real_world$Demand), 2)
)
final_period <- final_period %>%
  mutate(Capacity = c(3150, round(2.11 * 1050, 0)) - Demand)

plot_ly(
  final_period,
  x = ~PCF, y = ~Demand, type = "bar", name = "Demand", 
  marker = list(color = "ff7f0e")
) %>%
  add_trace(
    x = ~PCF, y = ~Capacity, type = "bar", name = "Capacity",
    marker = list(color = "1f77b4")
  ) %>%
  layout(
    title = "Final Model Representation",
    xaxis = list(title = ""),
    yaxis = list(title = "Veh per Period"),
    barmode = "stack"
  )

final_period %>%
  mutate(
    Capacity = Demand + Capacity,
    `% of Capacity` = round(Demand / Capacity * 100, 2)
  ) %>%
  kable() %>%
  kable_styling(full_width = FALSE)
```

## Calculation from survey
The PCFs for the model can be calculated from the trips in motion table using
the same approach. The results are shown in the table below.

```{r calc_pcf}
breaks <- c(am_peak_from, am_peak_to, pm_peak_from, pm_peak_to)
breaks <- as.numeric(breaks) * 1440
labels <- c("AM", "MD", "PM")

# an = function(n, len) c(seq.int(n), rep(n, len-n))
n = an(60 / bin_size, nrow(timTbl))

pcf_tbl <- timTbl %>%
  group_by(tod = cut(time, breaks = breaks, labels = labels)) %>%
  ungroup() %>%
  mutate(
    tod = as.character(tod),
    tod = ifelse(is.na(tod), "NT", tod),
    # Use factors to set display order in table
    tod = factor(
      tod,
      levels = c("AM", "MD", "PM", "NT"),
      ordered = TRUE
      )
    ) %>%
  group_by(tod) %>%
  mutate(
    # all_trips_in_hour = rollapplyr(AllTrips, 60 / bin_size, sum, partial = TRUE)
    all_trips_in_hour = frollsum(AllTrips, an(60 / bin_size, n()), na.rm = TRUE, adaptive = TRUE)
  ) %>%
  summarise(
    PHF = max(all_trips_in_hour) / sum(AllTrips),
    PCF = 1 / PHF
  ) %>%
  mutate(
    From = c(am_peak_from, am_peak_to, pm_peak_from, pm_peak_to),
    To = c(am_peak_to, pm_peak_from, pm_peak_to, am_peak_from),
    Length = ifelse(
      tod != "NT",
      (To - From) * 1440 / 60,
      (1 - as.numeric(From) + as.numeric(To)) * 1440 / 60
    )
  ) %>%
  # Reorder and rename for display
  select(Period = tod, From, To, Length, PHF, PCF)

total <- round(sum(pcf_tbl$PCF), 2)

pcf_tbl %>%
  rename(`Length (hrs)` = Length, `Model Capacity (hrs)` = PCF) %>%
  kable(digits = 2) %>%
  kable_styling(full_width = FALSE)
```

A [general
guideline](https://www.fhwa.dot.gov/planning/tmip/publications/other_reports/validation_and_reasonableness_2010/ch08.cfm)
to follow suggests the model should provide 8-12 hours of capacity per day. The
PCFs calculated directly from the survey data provide `r total`. This is slightly
higher but acceptable.


```{r write param tables, eval=FALSE}
# All writing of parameters files should be in this chunk, and will not
# execute when knitting. Run them manually only when you want to update the
# model.

# TOD factors
write_csv(tod_factors, "data/output/tod/time_of_day_factors.csv")

# Directionality factors. HBU gets appended by University.Rmd.
write_csv(pa_tbl , "../master/directionality/directionality_factors.csv")

# Occupancy factors. HBU gets appended by University.Rmd
write_csv(occ_tbl , "../master/directionality/veh_occ_factors.csv")

# Period capacity table
write_csv(
  pcf_tbl %>%
    select(TOD = Period, Value = PCF) %>%
    mutate(
      Description = paste("Hours of capacity in", TOD, "period", sep = " "),
      Value = round(Value, 3)
    ),
  "data/output/tod/period_capacity_factors.csv"
)
```
