---
title: "Untitled"
author: "name"
date: "6/4/2021"
output: html_document
---

```{r setup, include=FALSE}
source("R/university_surveysummary.R")


knitr::opts_chunk$set(echo = FALSE)
options(dplyr.summarise.inform = FALSE)
options(scipen = 999)

```

## Trip Length Distribution

Trip length distribution by purpose from NCSU survey. Distance is zone to zone distance from auto skim.

```{r}
zonetozonedistance_allpurposes_histogram + labs(title = "Trip Length Distribution (All purposes,except UC1)", 
                                              subtitle="for off-campus students (On_campus== 0) and on-campus students (On_campus == 1) students", 
                                              caption = ("Source: NCSU survey (unweighted)"))
                                              
zonetozonedistance_UC1_histogram + labs(title = "Trip Length Distribution (UC1 Trips)", 
                                        subtitle="for off-campus students (On_campus== 0) and on-campus students (On_campus == 1) students", 
                                        caption = ("Source: NCSU survey (unweighted)"))
```

## Mode Split

Mode split by Trip Purpose from NCSU survey
### On-Campus Students
```{r}
trips_bypurpose_oncampus_pct_df %>%
  kable(digits = 1) %>%
kable_styling(full_width = FALSE)

trips_bypurpose_oncampus_plot + labs(title = "Mode Split by Purpose", 
                                     subtitle="On-campus students", 
                                     caption = ("Source: NCSU survey (unweighted)"))

```

### Off-Campus Students
```{r}
trips_bypurpose_offcampus_pct %>%
  kable(digits = 1) %>%
kable_styling(full_width = FALSE)

trips_bypurpose_offcampus_plot + labs(title = "Mode Split by Purpose", 
                                     subtitle="Off-campus students", 
                                     caption = ("Source: NCSU survey (unweighted)"))

```


Although it is possible that students without a car drove, trips where students who did not have a car drove have been removed. 

###On-Campus Students with and without car
```{r, echo=FALSE}
trips_bypurpose_oncampus_caravail_pct2 %>%
kable(digits = 1) %>%
kable_styling(full_width = FALSE)
```

###Off-Campus Students with and without car
```{r}
trips_bypurpose_offcampus_caravail_pct2 %>%
kable(digits = 1) %>%
kable_styling(full_width = FALSE)
```

