---
title: "University"
author: "Caliper Corporation"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# renv::restore()

source("R/university_productions.R")
source("R/university_attractions.R")
# source("R/university_surveyTLD2.R")
# source("R/university_surveyMC.R")

knitr::opts_chunk$set(echo = FALSE)
options(dplyr.summarise.inform = FALSE)
options(scipen = 999, digits = 3)
```

## Introduction  

The University Model captures the on- and off-campus travel behavior of students of the Triangle Region's four large universities:  North Carolina State University (NCSU), The University of North Carolina at Chapel Hill (UNC), Duke University (Duke) and North Carolina Central University (NCCU). In 2016, the four universities had a combined enrollment of `r enrollment_total`, with `r enrollment_NCSU` students at NCSU, `r enrollment_UNC` at UNC, `r enrollment_Duke` at Duke and `r enrollment_NCCU` at NCCU.  

```{r enrollment,echo=FALSE}
enrollment_NCSU <- socioecon2_df %>% 
  filter(!is.na(StudGQ_NCSU),!is.na(StudOff_NCSU)) %>% 
  summarize(total=sum(StudGQ_NCSU,StudOff_NCSU)) %>% 
  pull(total)

enrollment_UNC <- socioecon2_df %>% 
  filter(!is.na(StudGQ_UNC),!is.na(StudOff_UNC)) %>% 
  summarize(total=sum(StudGQ_UNC,StudOff_UNC)) %>% 
  pull(total)

enrollment_Duke <- socioecon2_df %>% 
  filter(!is.na(StudGQ_DUKE),!is.na(StudOff_DUKE)) %>% 
  summarize(total=sum(StudGQ_DUKE,StudOff_DUKE)) %>%
  pull(total)

enrollment_NCCU <- socioecon2_df %>% 
  filter(!is.na(StudGQ_NCCU),!is.na(StudOff_NCCU)) %>% 
  summarize(total=sum(StudGQ_NCCU,StudOff_NCCU)) %>%
  pull(total)

enrollment_total <- enrollment_NCSU + enrollment_UNC + enrollment_Duke + enrollment_NCCU
```


On- and off-campus students are expected to have different travel behavior. On-campus students are expected to make more trips and most of those trips are expected to start and end on campus. On-campus students are more likely to be undergraduate students who are enrolled at the university full-time while graduate students are more likely to reside off-campus.

The model is segmented by on-campus and off-campus student and by the following trip purposes:

* Home-Based-Campus (UHC)  
*	Home-Based-Other (UHO)  
*	Campus-Based-Other (UCO)  
*	On-Campus (UC1)  
*	Inter-Campus (UCC)  
*	University student Other-Other (UOO)  

## TRIP GENERATION  

### TRIP PRODUCTION RATES  

#### Sources  

##### 2014 NCSU Survey  


```{r survey_overview, echo = FALSE}
Respondents_On_campus <- surveyRespondents_ResidenceClass_df %>%
  group_by(On_campus) %>%
  filter(On_campus==1) %>%
  summarize (sum=sum(n)) %>%
  pull()
Respondents_Off_campus <- surveyRespondents_ResidenceClass_df %>%
  group_by(On_campus) %>%
  filter(On_campus==0) %>%
  summarize (sum=sum(n)) %>%
  pull()
Respondent_Off_Campus_Weighted <- SurveyRespondent_Residence_Weighted_df %>%
  filter(On_campus==0)%>%
  pull()  
Respondent_On_Campus_Weighted <- SurveyRespondent_Residence_Weighted_df %>%
  filter(On_campus==1)%>%
  pull()

Survey_Total_Weighted_Trips <- Trip_subset_df %>% 
  filter(!is.na(Weight)) %>% 
  summarize(sum(Weight))
```

The trip production rates are developed based on a survey of NCSU students conducted by the Institute for Transportation Research and Education (ITRE) for the North Carolina Department of Transportation (NCDOT) in 2014. After removing respondents who did not finish the survey and respondents with a mismatch of more than 1 between the stated number of trips and the number of trips for which they provided a description, the total number of complete and valid samples consists of `r Respondents_On_campus` on-campus students and `r Respondents_Off_campus` and off-campus students. The breakdown by graduate and undergraduate students is presented in the table below. These respondents reported a combined `r nrow(Trip_subset_df)` trips.

```{r survey overview2,echo=FALSE}
surveyRespondents_ResidenceClass_df %>%
  kable(digits = 3) %>%
  kable_styling(full_width = FALSE)
```


The NCSU survey dataset includes person weights which were developed by ITRE based on residence locations (on-campus vs. off-campus), credit hours (full-time vs. part-time) and class status (undergraduate vs. graduate).  A total of `r Respondent_On_Campus_Weighted` on_campus and `r Respondent_Off_Campus_Weighted` off-campus records were assigned a weight. Based on the survey, the NCSU student population made a total of `r Survey_Total_Weighted_Trips` trips in 2016.



##### 2012 Virginia Surveys   
The trip production rates developed based on the NCSU survey were verified using trip rates based on surveys of students from 4 universities in Virginia conducted in 2009 and repeated for two of the universities in 2010, as summarized in Khattak et al. (2012). The first round of surveys obtained a sample size of 2784. The sample size of the second round was 2596.

#### Segmentation   

The NCSU survey data was used to develop trip production rates for on-campus and off-campus students for 5 of the model's trip purposes.  

* Home-Based-Campus (UHC): P = Home  
* Home-Based-Other (UHO): P = Home  
* Campus-Based-Other(UCO): P = Campus  
* On-Campus (UC1): P = Campus  
* University student Other-Other (UOO): P = Origin 
* Campus to Campus (UCC): P = Origin


#### Exploratory Analysis 

Figure 1 shows the average number of trips per student per day for each of the 5 trip purposes for students residing on-campus and off-campus, based on the raw NCSU survey data. 
The distribution is skewed to the right, with most students making few trips and a few students making many trips per day.There appears to be more variability in the number of trips reported by on-campus students, especially in UC1 and UCH, which are trips that both start and end on campus for on-campus students. 


```{r histogram,echo=FALSE}
Triprates_histogram<-Productions_bymode_df %>% 
  group_by(Person_ID,
           Trip_Purpose, 
           On_campus)%>%
  summarize(Trips_per_student=sum(Trips),.groups = "drop")%>%
  group_by(On_campus) %>% 
  ggplot(aes(Trips_per_student, fill=Trip_Purpose)) + 
  geom_histogram(bins=20) + 
  facet_grid(Trip_Purpose~On_campus)

Triprates_histogram + labs(title = "Figure 1 - Trips per student by purpose", 
                          subtitle="for off-campus (On_campus==0) and on-campus (On_campus==1) students", 
                          caption = ("Source: NCSU survey (unweighted)"))
```

To develop production rates, two methods were explored: cross-classification and regression. Explanatory variables in regression models tested included student characteristics such as class status (undergraduate and graduate), on and/or off-campus employment, credit hours (full-time or part-time), car availability as these were expected and/or shown by other research to affect university trip production rates.  Because these explanatory variables are not part of the future year socioeconomic datasets, application of these models would not be feasible.  The cross-classification findings are discussed below.

#### Cross-classification  

As part of the cross-classification approach, daily trips per student reported in the NCSU survey were classified based on the following criteria:  
1) place of residence: on-campus or off-campus  
2) trip purpose: UHC, UHO, UCO, UC1 or UOO   
3) mode: walk, bike, bus (includes public bus and shuttle), drive, carpool and other  

Production rates were developed based on raw data and based on weighted data. The rates based on the weighted data are very similar to the rates based on the raw data.  

Based on the weighted data, the average daily trips rate is `r w_avg_trips_df[,"avg_trips"]`.  On-campus students make an average of `r w_avg_trips_byresidence_df[2,"avg_trips"]` trips per day while off- campus students make an average of `r w_avg_trips_byresidence_df[1,"avg_trips"]` trips per day. Rates by trip purpose show that on-campus students make more on-campus trips than off-campus students but make fewer off-campus trips.


Comparison to Khattak et al (201

Consistent with our findings, Khattak et al found that students who reside on-campus make more trips (especially on-campus trips) than those who reside off-campus and that they walk more and drive less than off-campus students. They estimate production trip rates that were somewhat lower than the production rates selected for this study. During first round of surveys at ODU, VT, UVA and VCU, daily trip rates ranged from 4.4 to 4.9.  For the second round of surveys, the survey instruments were revised to reduce the survey response burden and a larger sample size was obtained. Based on the second round, daily trip rates were 5.3 at ODU and 5.6 at VT. 

```{r,echo=FALSE}


P_rates_df %>%
  kable(digits = 3) %>%
  kable_styling(full_width = FALSE)
```
##### UOO Trip Rates by Mode  
Trip production rates by mode will be used to estimate a trip production rate for UOO trips based on the relationship between trip production by mode for UOO trips and for UHO and UCO trips.  SCOPE: "The UOO trips should be generated separately by mode as a function of (or rate per) UHO and UCO off-campus trip ends by mode and optionally a nearby accessibility variable (provided by Caliper).  UOO trip generation and distribution models will therefore be implemented to run at the end of the university model after the other trip purposes"

```{r test, echo=FALSE}

kable(avg_trips_UHOUCO_df[,avg_trips_UHOUCO_col],caption ="UHO and UCO Trip Rates by Mode",col.names=headers) 

kable(avg_trips_UOO_df[,avg_trips_UOO_col],caption ="UOO Trip Rates by Mode",col.names=headers)

kable(P_rates_ratioUOOtoUHOUCO_df, caption = "Ratio UOO Production Rate to UHOUCO Production Rate", col.names=c("car","non-car"))
```

#####check if group quarters are always on-campus; if not only include non-campus group quarters 
Trip productions by TAZ were estimated by applying the trip production rates as follows:  
On-Campus Students:  
* UC1 Productions =  
Share of campus buildings in TAZ * enrollment *  UC1 On-campus production rate  
* UCO Productions=  
Share of campus buildings in TAZ * enrollment *  UCO On-campus production rate  
* UHC Productions =   
Number of students in student group quarters in *   UHC On-campus production rate  
* UHO Productions =  
Number of students in student group quarters *   UHO On-campus production rate  
* UCC Productions =  
Number of students in student group quarters *   UCC On-campus production rate

Off-Campus Students:  
* UC1 Productions =  
Share of campus buildings in TAZ * enrollment *  UC1 Off-campus production rate  
* UCO Productions =  
Share of campus buildings in TAZ * enrollment *  UCO Off-campus production rate  
* UHC Productions =  
Number of students in off-campus housing  *   UHC Off-campus production rate  
* UHO Productions =   
Number of students in off-campus housing *   UHO Off-campus production rate  
* UCC Productions =  
Number of students in off-campus housing *   UCC Off-campus production rate


```{r}
Summary_Productions_df %>%
  kable(digits = 0) %>%
  kable_styling(full_width = FALSE)
```


###  TRIP ATTRACTION RATES 

#### Sources

##### Off-campus attractions
As the trip production rates, trip attraction rates for off-campus attractions (UHO and UCO) were developed based on a survey of NCSU students conducted by ITRE for NCDOT in 2014. For UHO trips by off-campus students, both trips ends are off-campus. The number of trip attractions is estimated as a function of the land use of the zone.  For UCO trips and for UHO trips by on-campus students, the production end is the campus. Therefore, in addition to land use variables of the attraction zone, the attraction model for these segments also considers distance between the zone and campus.

##### On-campus attractions
For on-campus attractions (UHC,UC1), the attractions are distributed among campus zones proportional to the distribution of the building square feet. 

#### Exploratory Analysis  

##### Correlations  
Correlations between the university attractions and land use characteristics at the TAZ level are presented below. For trip attractions by on-campus students, the  off-campus student population and the number of retail jobs have the highest correlation coefficients.  For trips by off-campus students, off-campus student population has the highest correlation coefficient.  
```{r,echo=FALSE}
correlations_df %>% 
 select(term,AllStudents_Trips, OnCampusStudents_Trips, OffCampusStudents_Trips) %>%
  kable(digits = 3) %>%
  kable_styling(full_width = FALSE)

```

#### Regression Models
The regression models were estimated for each purpose using a dataset that includes all TAZs for which the survey included trip attractions, of any purpose.  The TAZs that do not  have any trip attractions in the survey  were excluded from the dataset used for model estimation.
```{r, echo=FALSE}
print_model <- function(model) {
  df <- tibble(variable = colnames(model$model),
               estimate = model$coefficients,
               statistic = coef(summary(model))[, "t value"],
               p.value = coef(summary(model))[, "Pr(>|t|)"]) %>%
    mutate(variable = if_else(variable == "y", "_intercept_", variable)) %>%
    arrange(variable) %>%
    bind_rows(., tibble(variable = c("Adjusted R-squared"),
                        estimate = summary(model)$adj.r.squared,
                        statistic = as.double(NA),
                        p.value = as.double(NA))) %>%
    kable(digits = 4) %>%
    kable_styling(full_width = FALSE)
  
  return(df)
  
}
```

##### UCO trips by On-campus students

The  model explains the number of trip attractions at the TAZ level based on the number of the students residing in the TAZ and the number of retail jobs in the TAZ. The coefficients have the expected signs. 
```{r, echo=FALSE}
print_model(Model_OnCampusUCO_2)

```

##### UHO trips by On-campus students
As UCO trips by on-campus students, UHO trips by on-campus students are trips between the campus and off-campus location. The model explains the number of trips based on the number of off-campus students residing in the TAZ, the number of retail jobs and the distance to campus. The distance is calculated as the average of the distances between the centroid of the TAZ  and each of the centroids of the different  NCSU campus TAZs.  All coefficients have the expected sign.  As expected, trip attractions decrease as the distance increases and trip attractions increase as student population and retail jobs increase. 

```{r,echo=FALSE}
print_model(Model_Oncampus_UHO_2)

```

##### UcO trips by Off-campus students
As UCO and UHO trips by on-campus students, UCO trips by off-campus students are trips between the campus and off-campus location. The model explains UCO trips by off-campus students based on distance to campus, off- campus student population and  number of retail jobs.
```{r,echo=FALSE}
print_model(Model_Offcampus_UCO_2)

```

#####  UHO Trips by Off-campus students  
UHO trips by off-campus students have two off-campus trip ends. The model explains UHO trips by off-campus students based on the off- campus student population and the number of retail jobs.
```{r, echo=FALSE}
print_model(Model_Offcampus_UHO_2)


```
## TRIP DISTRIBUTION  
### Trip length distribution from the NCSU survey
```{r,echo=FALSE}
zonetozonedistance_bypurpose_histogram + labs(title = "Trip Length Distribution (All purposes,except UC1)", 
                                              subtitle="for off-campus students (On_campus== 0) and on-campus students (On_campus == 1) students", 
                                              caption = ("Source: NCSU survey (unweighted)"))


zonetozonedistance_UC1_histogram + labs(title = "Trip Length Distribution (UC1 Trips)", 
                                        subtitle="for off-campus students (On_campus== 0) and on-campus students (On_campus == 1) students", 
                                        caption = ("Source: NCSU survey (unweighted)"))
```

## MODE CHOICE
### Mode split from the NCSU survey
#### On-campus students mode split
```{r,echo=FALSE}
modesplit_bypurpose_oncampus_2_df 
```
#### Off-campus students mode split
```{r,echo=FALSE}
modesplit_bypurpose_offcampus_2_df
```
### Mode split calibration targets
#### On-campus students mode split calibration targets
```{r,echo=FALSE}
calibrationtargets_bypurpose_oncampus_df
```
#### Off-campus students mode split calibration targets
```{r,echo=FALSE}
calibrationtargets_bypurpose_offcampus_df
```

